---
title: "carbon_RQ"
output: html_notebook
---

# Analysis per Research Question

## RQ1

```{r}


analysis_sample <- analysis_sample %>%
  mutate(
    cpi10   = carbon_pricing_index/10,             # per +10 CPI pts
    t       = year,                                 # trend index if needed
    lgdppc_c= scale(log_gdp_pc, scale = FALSE),
    ind_c   = scale(industry_share, scale = FALSE),
    ren_c   = scale(renewable_share, scale = FALSE)
  )

# Baseline TWFE (2-way clustered)
m1 <- feols(log_co2 ~ cpi10 + lgdppc_c + ind_c + ren_c | country + year,
            cluster = ~ country + year, data = analysis_sample)

# Driscoll–Kraay SEs
m2 <- feols(log_co2 ~ cpi10 + lgdppc_c + ind_c + ren_c | country + year,
            data = analysis_sample, panel.id = ~ country + year,
            vcov = vcov_DK(~ year, lag = 3))

etable(m1, m2, se.below = TRUE, headers = c("TWFE 2w-cluster","TWFE DK"))

```
```{r}
sun_df <- analysis_sample %>%
  group_by(country) %>%
  mutate(adopt_year = ifelse(any(carbon_pricing_index>0, na.rm=TRUE),
                             min(year[carbon_pricing_index>0], na.rm=TRUE), 0L)) %>%
  ungroup()

m_sun <- feols(log_co2 ~ log_gdp_pc + industry_share + renewable_share +
                 sunab(adopt_year, year) | country + year,
               data = sun_df, cluster = ~ country + year)

iplot(m_sun, ref.line = 0, main = "Event Study: First CPI>0")

```

## RQ2 (Heterogeneity by carbon intensity / renewables)

```{r}
library(dplyr)
library(fixest)

# --- 0) Ensure panel keys exist and are unique --------------------------------
if (!"country" %in% names(merged_final) && "iso3" %in% names(merged_final)) {
  merged_final <- dplyr::rename(merged_final, country = iso3)
}
stopifnot(!anyDuplicated(merged_final[c("country","year")]))

# --- 1) Renewables: normalize column, add lag (we will use the lag in regressions)
# If you have renewable_share_pc (0-100), convert to fraction; otherwise keep renewable_share as-is
merged_final <- merged_final %>%
  mutate(
    renewable_share = dplyr::case_when(
      !is.na(renewable_share) ~ renewable_share,
      !is.na(renewable_share_pc) ~ renewable_share_pc / 100,
      TRUE ~ NA_real_
    )
  ) %>%
  group_by(country) %>%
  arrange(year, .by_group = TRUE) %>%
  mutate(renewable_share_l1 = lag(renewable_share, 1)) %>%
  ungroup()

# --- 2) Build analysis sample (renewables start ~2004; CO2 available to 2023) ---
analysis_sample <- merged_final %>%
  filter(year >= 2004, year <= 2023) %>%
  mutate(
    cpi10    = carbon_pricing_index / 10,         # effect per +10 CPI points
    lgdppc_c = scale(log_gdp_pc, scale = FALSE),  # mean-centered controls
    ind_c    = scale(industry_share, scale = FALSE),
    ren_c    = scale(renewable_share, scale = FALSE),
    ren_c_l1 = scale(renewable_share_l1, scale = FALSE),
    t        = year
  ) %>%
  tidyr::drop_na(log_co2, cpi10, lgdppc_c, ind_c, ren_c_l1)  # use lagged renewables to reduce endog.

stopifnot(!anyDuplicated(analysis_sample[c("country","year")]))

# --- 3) TWFE models ------------------------------------------------------------
# 3a) Baseline (two-way clustered)
m1 <- feols(
  log_co2 ~ cpi10 + lgdppc_c + ind_c + ren_c_l1 | country + year,
  cluster = ~ country + year,
  data = analysis_sample
)

# 3b) Driscoll–Kraay (needs panel.id + DK time index)
m2 <- feols(
  log_co2 ~ cpi10 + lgdppc_c + ind_c + ren_c_l1 | country + year,
  data     = analysis_sample,
  panel.id = ~ country + year,
  vcov     = vcov_DK(~ year, lag = 3)
)

# 3c) Country-specific linear trends (robustness)
m3 <- feols(
  log_co2 ~ cpi10 + lgdppc_c + ind_c + ren_c_l1 | country[t] + year,
  cluster  = ~ country + year,
  data     = analysis_sample
)

# 3d) Dynamics: CPI lags 0–2 with DK SEs
m4 <- feols(
  log_co2 ~ l(cpi10, 0:2) + lgdppc_c + ind_c + ren_c_l1 | country + year,
  data     = analysis_sample,
  panel.id = ~ country + year,
  vcov     = vcov_DK(~ year, lag = 3)
)

etable(m1, m2, m3, m4, se.below = TRUE,
       headers = c("TWFE 2w-cluster","TWFE DK","TWFE+trends","TWFE+lags"))

# Cumulative CPI effect over 0–2 years (robust to naming quirks)
lag_names <- grep("^l\\(cpi10,\\d\\)$", names(coef(m4)), value = TRUE)
b <- coef(m4)[lag_names]; V <- vcov(m4)[lag_names, lag_names, drop = FALSE]
cum_est <- sum(b); cum_se <- sqrt(t(rep(1, length(b))) %*% V %*% rep(1, length(b)))
cum_z <- as.numeric(cum_est / cum_se); cum_p <- 2 * pnorm(-abs(cum_z))
c(cum_est = cum_est, se = cum_se, z = cum_z, p = cum_p)

```

## RQ3 Dynamics Lags

```{r}
m_dyn <- feols(log_co2 ~ l(cpi10,0:2) + lgdppc_c + ind_c + ren_c | country + year,
               data = analysis_sample, panel.id = ~ country + year,
               vcov = vcov_DK(~ year, lag = 3))

# Cumulative CPI effect over 0–2 years (robust to naming issues)
lag_names <- grep("^l\\(cpi10,\\d\\)$", names(coef(m_dyn)), value = TRUE)
b <- coef(m_dyn)[lag_names]; V <- vcov(m_dyn)[lag_names, lag_names]
est <- sum(b); se <- sqrt(matrix(1,1,3) %*% V %*% matrix(1,3,1)); z <- est/se; p <- 2*pnorm(-abs(z))
c(cum_est = est, se = se, z = z, p = p)

```

```{r}
# Baseline intensity pre-2004
base_int <- merged_final %>%
  filter(year >= 2000, year <= 2003) %>%
  mutate(intensity = co2 / gdp) %>%
  group_by(country) %>%
  summarise(base_int = median(intensity, na.rm = TRUE), .groups = "drop")

analysis_sample <- analysis_sample %>%
  left_join(base_int, by = "country") %>%
  mutate(base_int_c = scale(base_int, scale = FALSE))

m_het_intensity <- feols(
  log_co2 ~ cpi10*base_int_c + lgdppc_c + ind_c + ren_c_l1 | country + year,
  cluster = ~ country + year, data = analysis_sample
)
etable(m_het_intensity, se.below = TRUE)

```

## RQ 4 5

```{r}
# Lag EPS to reduce simultaneity
analysis_sample <- analysis_sample %>%
  group_by(country) %>% arrange(year, .by_group = TRUE) %>%
  mutate(eps_l1 = lag(oecd_value, 1)) %>%
  ungroup() %>%
  mutate(eps_c = scale(eps_l1, scale = FALSE))

m_eps <- feols(
  log_co2 ~ cpi10*eps_c + lgdppc_c + ind_c + ren_c_l1 | country + year,
  cluster = ~ country + year, data = analysis_sample
)

# ETS phases (interactions only; main effects collinear with year FE)
analysis_sample <- analysis_sample %>%
  mutate(phaseII  = (year>=2008 & year<=2012),
         phaseIII = (year>=2013 & year<=2020),
         phaseIV  = (year>=2021 & year<=2023))

m_phase <- feols(
  log_co2 ~ cpi10:phaseII + cpi10:phaseIII + cpi10:phaseIV +
            lgdppc_c + ind_c + ren_c_l1 | country + year,
  cluster = ~ country + year, data = analysis_sample
)
# Is Phase IV stronger than earlier phases?
wald(m_phase, "cpi10:phaseIV = cpi10:phaseIII")
wald(m_phase, "cpi10:phaseIV = cpi10:phaseII")

```

# Analysis per Research Question

## RQ1 / H1 Average causal impact

```{r}

# Main (per-capita outcome)
m1 <- feols(log_co2_pc ~ lcpi + ren_c + lgdppc_c + ind_c | iso3 + year,
            cluster=~iso3, data=analysis_sample)

# With country-specific linear trends (absorbs slow paths)
m1_tr_i <- feols(log_co2_pc ~ lcpi + ren_c + lgdppc_c + ind_c + i(iso3, t) | iso3 + year,
  cluster = ~ iso3, data = analysis_sample
)

# Totals as robustness, controlling for size (STIRPAT-style)
m1_tot <- feols(log_co2 ~ lcpi + log_pop + log_gdp_pc + ren_c + ind_c | iso3 + year,
                cluster=~iso3, data=analysis_sample)
etable(m1, m1_tr_i, m1_tot, se.below=TRUE,
       headers=c("Per-capita","Per-capita+trends","Totals"))

```
## RQ 3 / H3 Dynamics
```{r}
# Main Specification: Per-capita emissions with lags
m_main <- feols(
  log_co2_pc ~ l(cpi10, 0:2) + lgdppc_c + ind_c + ren_c | country + year,
  data     = analysis_sample,
  panel.id = ~ country + year,
  vcov     = vcov_DK(~ year, lag = 3)
)

# Robustness Check: Total emissions with size controls and lags
m_robust <- feols(
  log_co2 ~ l(cpi10, 0:2) + log_pop + log_gdp_pc + ind_c + ren_c | country + year,
  data    = analysis_sample,
  panel.id = ~ country + year,
  vcov     = vcov_DK(~ year, lag = 3)
)

etable(m_main, m_robust, se.below = TRUE,
       headers = c("Per-capita (main)","Totals (robustness)"))
```

Two-way FE with Driscoll–Kraay SEs shows no contemporaneous CPI effect, but a one-year-lagged reduction of ~1.5% (per-capita) and ~2.1% (totals) per +10 €/t increase in the carbon price. The 2-year lag is small and insignificant, consistent with adjustment occurring within roughly a year. We’re stacking level lags of a highly persistent policy (CPI), so the t+1 effect shows the first round of adjustment, while t+2 gets diluted by (i) multicollinearity with newer CPI levels, (ii) front-loaded abatement, and (iii) annual measurement/budget cycles.

```{r}
# Formally testing cumulative effect over 0-2 years
# Wald test for m_main, m_robust

cum_wald <- function(mod, coefs = c("cpi10","l(cpi10, 1)","l(cpi10, 2)")) {
  nm  <- names(coef(mod))
  idx <- match(coefs, nm)
  if (any(is.na(idx))) stop("Missing coefs: ", paste(coefs[is.na(idx)], collapse=", "))

  b <- coef(mod)[idx]
  V <- vcov(mod)[idx, idx, drop = FALSE]  # uses the model's DK vcov
  L <- matrix(1, nrow = 1, ncol = length(idx))

  est <- as.numeric(L %*% b)
  se  <- sqrt(as.numeric(L %*% V %*% t(L)))
  z   <- est / se
  p   <- 2 * pnorm(abs(z), lower.tail = FALSE)
  ci  <- est + c(-1, 1) * 1.96 * se
  list(est = est, se = se, z = z, p = p, ci = ci)
}

res_main   <- cum_wald(m_main)
res_robust <- cum_wald(m_robust)
res_main
res_robust

```
Two-way FE with Driscoll–Kraay SEs shows no contemporaneous CPI effect, but a cumulative two-year reduction of ≈1.0% (per-capita) and ≈1.9% (totals) for a one-time +10 €/t increase in carbon price (95% CIs [−1.90, −0.17] and [−2.82, −0.96]). The decline is concentrated at t+1.

```{r}

# --- assume you already have m_main and m_robust (with DK SEs, lags 0..2) ---

# helper to compute cumulative (0–2y): cpi10 + l1 + l2
cum_02 <- function(mod){
  nm <- names(coef(mod))
  want <- c("cpi10","l(cpi10, 1)","l(cpi10, 2)")
  idx <- match(want, nm)
  b <- coef(mod)[idx]
  V <- vcov(mod)[idx, idx, drop = FALSE]
  L <- matrix(1, 1, 3)

  est <- as.numeric(L %*% b)
  se  <- sqrt(as.numeric(L %*% V %*% t(L)))
  z   <- est / se
  p   <- 2 * pnorm(abs(z), lower.tail = FALSE)
  stars <- ifelse(p < .001,"***", ifelse(p < .01,"**", ifelse(p < .05,"*",
                   ifelse(p < .1,".",""))))
  list(est=est, se=se, p=p, stars=stars)
}

r_main   <- cum_02(m_main)
r_robust <- cum_02(m_robust)

# pretty print cumulative cells, e.g., "-0.016** (0.0048)"
fmt <- function(x) formatC(x, format="f", digits=3)
cum_row <- data.frame(
  term = "Cumulative CPI (0–2y)",
  `Per-capita (main)`   = glue("{fmt(r_main$est)}{r_main$stars}\n({fmt(r_main$se)})"),
  `Totals (robustness)` = glue("{fmt(r_robust$est)}{r_robust$stars}\n({fmt(r_robust$se)})")
)

# map coefficient names for publication
coef_map <- c(
  "cpi10"        = "CPI (t)",
  "l(cpi10, 1)"  = "CPI (t−1)",
  "l(cpi10, 2)"  = "CPI (t−2)",
  "log_pop"      = "log(Population)",
  "log_gdp_pc"   = "log(GDP per capita)",
  "lgdppc_c"     = "log(GDPpc) (demeaned)",
  "ind_c"        = "Industry share (demeaned)",
  "ren_c"        = "Renewables share (demeaned)"
)

# which GOF rows to show and how to label them
gof_map <- tribble(
  ~raw,               ~clean,                 ~fmt,
  "nobs",             "Observations",         0,
  "r.squared",        "R^2",                  3,
  "within.r.squared", "Within R^2",           3
)

# Add fixed effects indicators as custom rows
fe_rows <- data.frame(
  term = c("Country FE", "Year FE"),
  `Per-capita (main)`   = c("Yes","Yes"),
  `Totals (robustness)` = c("Yes","Yes")
)

# build the table
msummary(
  list("Per-capita (main)" = m_main, "Totals (robustness)" = m_robust),
  coef_map  = coef_map,
  gof_map   = gof_map,
  add_rows  = rbind(cum_row, fe_rows),
  statistic = "({std.error})",
  stars     = c('*' = .05, '**' = .01, '***' = .001),
  notes = list(
    "Two-way fixed effects at country and year levels.",
    "Standard errors are Driscoll–Kraay with 3 lags, clustered by year dimension.",
    "Dependent variable is log CO₂ per-capita (col. 1) and log total CO₂ (col. 2).",
    "CPI is scaled so that a +10 €/t price change equals +1 in cpi10.",
    "Cumulative CPI (0–2y) equals CPI(t) + CPI(t−1) + CPI(t−2)."
  ),
  output = "latex"
)

```

Checking lag collinearity:
```{r}

tmp <- analysis_sample %>%
  arrange(country, year) %>%
  group_by(country) %>%
  mutate(cpi_l1 = lag(cpi10), cpi_l2 = lag(cpi10, 2)) %>%
  ungroup()

with(tmp, c(
  cor01 = cor(cpi10, cpi_l1, use="pair"),
  cor02 = cor(cpi10, cpi_l2, use="pair"),
  cor12 = cor(cpi_l1, cpi_l2, use="pair")
))

```
Problem: Multicollinearity of Lags
The lagged values of CPI are highly correlated with each other (e.g., cor(cpi10, lag(cpi10)) ≈ 0.92).

The Solution: Almon Polynomial Distributed Lags (PDL)
The Almon lag (or PDL) model is designed specifically to solve this problem. Instead of estimating a separate coefficient for each lag, it imposes a structure, assuming the coefficients of the lags lie along a smooth polynomial curve (in your case, a 2nd-degree polynomial, which is a parabola).

This approach overcomes multicollinearity by estimating the shape of the lag effect rather than each individual point, resulting in more stable and efficient estimates.
```{r}

analysis_sample$year <- as.integer(analysis_sample$year)

# sanity: no duplicate country-year rows?
any(duplicated(analysis_sample[c("country","year")]))

m_l3 <- feols(
  log_co2_pc ~ l(cpi10, 0:3) + lgdppc_c + ind_c + ren_c | country + year,
  data      = analysis_sample,
  panel.id  = ~ country + year,           
  vcov      = vcov_DK(~ year, lag = 3)
)
etable(m_l3, se.below = TRUE)

cum_wald(m_l3, coefs = c("cpi10","l(cpi10, 1)","l(cpi10, 2)","l(cpi10, 3)"))

```


```{r does not work}
#Smoothed distributed lag (Almon PDL): 
feols(log_co2_pc ~ almon(cpi10, 0:3, 2) + lgdppc_c + ind_c + ren_c | country + year,
      data = analysis_sample,
      panel.id = ~ country + year, 
      vcov = vcov_DK(~year, lag = 3))

m_almon <- feols(
  log_co2_pc ~ almon(cpi10, 0:3, 2) + lgdppc_c + ind_c + ren_c | country + year,
  data     = analysis_sample,
  panel.id = ~ country + year,
  vcov     = vcov_DK(~ year, lag = 3)
)

as_lp <- analysis_sample |>
  dplyr::arrange(country, year) |>
  dplyr::group_by(country) |>
  dplyr::mutate(dcpi = cpi10 - dplyr::lag(cpi10)) |>
  dplyr::ungroup()

run_lp <- function(h) feols(flead(log_co2_pc, h) ~ dcpi + lgdppc_c + ind_c + ren_c |
                            country + year, data = as_lp, vcov = vcov_DK(~year, lag = 3))
etable(lapply(0:3, run_lp), headers = paste0("h=",0:3), se.below=TRUE)
etable(m_l3, m_almon, se.below=TRUE)

```


```{r}


# Build 0..3 lags and degree-2 PDL basis
as_pdl <- analysis_sample %>%
  arrange(country, year) %>%
  group_by(country) %>%
  mutate(
    cpi_l0 = cpi10,
    cpi_l1 = lag(cpi10, 1),
    cpi_l2 = lag(cpi10, 2),
    cpi_l3 = lag(cpi10, 3),
    pdl0   = cpi_l0 + cpi_l1 + cpi_l2 + cpi_l3,                  # Σ k^0 L^k x
    pdl1   = 0*cpi_l0 + 1*cpi_l1 + 2*cpi_l2 + 3*cpi_l3,          # Σ k^1 L^k x
    pdl2   = 0^2*cpi_l0 + 1^2*cpi_l1 + 2^2*cpi_l2 + 3^2*cpi_l3   # Σ k^2 L^k x
  ) %>%
  ungroup()

m_pdl <- feols(
  log_co2_pc ~ pdl0 + pdl1 + pdl2 + lgdppc_c + ind_c + ren_c | country + year,
  data = as_pdl,
  vcov = vcov_DK(~ year, lag = 3)
)

etable(m_pdl, se.below = TRUE)
```
```{r}
# Implied lag betas and their SEs
a <- coef(m_pdl)[c("pdl0","pdl1","pdl2")]
V <- vcov(m_pdl)[c("pdl0","pdl1","pdl2"), c("pdl0","pdl1","pdl2")]

k <- 0:3
b_hat <- as.numeric(a[1] + a[2]*k + a[3]*k^2)
J <- cbind(1, k, k^2)                            # db/da
se_b <- sqrt(diag(J %*% V %*% t(J)))
lag_df <- data.frame(lag=k, beta=b_hat, se=se_b,
                     lo=b_hat - 1.96*se_b, hi=b_hat + 1.96*se_b)

# Cumulative effects 0–2 and 0–3 with CIs
Sigma_b <- J %*% V %*% t(J)                      # cov(implied betas)
L02 <- matrix(c(1,1,1,0), 1) ; L03 <- matrix(c(1,1,1,1), 1)
est02 <- as.numeric(L02 %*% b_hat)
se02  <- sqrt(as.numeric(L02 %*% Sigma_b %*% t(L02)))
ci02  <- est02 + c(-1,1)*1.96*se02

est03 <- as.numeric(L03 %*% b_hat)
se03  <- sqrt(as.numeric(L03 %*% Sigma_b %*% t(L03)))
ci03  <- est03 + c(-1,1)*1.96*se03

list(lag_table = lag_df,
     cum_0_2 = c(est=est02, se=se02, lo=ci02[1], hi=ci02[2]),
     cum_0_3 = c(est=est03, se=se03, lo=ci03[1], hi=ci03[2]))

```


```{r}
# After running your model...
gamma <- coef(m_pdl)[c("pdl0", "pdl1", "pdl2")]

# Calculate the effect for each lag
beta_k <- c(
  "lag0" = gamma["pdl0"],
  "lag1" = gamma["pdl0"] + gamma["pdl1"]*1 + gamma["pdl2"]*1^2,
  "lag2" = gamma["pdl0"] + gamma["pdl1"]*2 + gamma["pdl2"]*2^2,
  "lag3" = gamma["pdl0"] + gamma["pdl1"]*3 + gamma["pdl2"]*3^2
)

print(beta_k)
```


```{r}
# 0) ensure clean panel keys
analysis_sample$year <- as.integer(analysis_sample$year)
stopifnot(!any(duplicated(analysis_sample[c("country","year")])))

# 1) Build the CPI "shock": +1 in dcpi == +10 €/t increase
as_lp <- analysis_sample %>%
  arrange(country, year) %>%
  group_by(country) %>%
  mutate(dcpi = cpi10 - lag(cpi10)) %>%
  ungroup()

# 2) Local projection runner: y_{t+h} on dcpi_t with FE and DK SEs
run_lp <- function(h) {
  dat_h <- as_lp %>%
    group_by(country) %>%
    mutate(y_h = dplyr::lead(log_co2_pc, h)) %>%
    ungroup()
  feols(
    y_h ~ dcpi + lgdppc_c + ind_c + ren_c | country + year,
    data = dat_h,
    vcov = vcov_DK(~ year, lag = 3)
  )
}

mods <- lapply(0:3, run_lp)
etable(mods, headers = paste0("h=", 0:3), se.below = TRUE)

```

1) Baseline distributed-lag table (Per-capita & Totals)

This reproduces your main two specs with CPI at t, t-1, t-2, and adds a custom cumulative (t..t+2) row.

```{r repetition of baseline}

# helper: cumulative effect & SE for selected coefficients
cum_wald <- function(mod, coefs) {
  nm  <- names(coef(mod))
  idx <- match(coefs, nm)
  b   <- coef(mod)[idx]
  V   <- vcov(mod)[idx, idx, drop = FALSE]
  L   <- matrix(1, nrow = 1, ncol = length(idx))
  est <- as.numeric(L %*% b)
  se  <- sqrt(as.numeric(L %*% V %*% t(L)))
  list(est = est, se = se)
}
fmt_cell <- function(est, se, digits = 3) sprintf("%.3f\n(%.3f)", est, se)

# coefficient labels
cm <- c(
  "cpi10"        = "CPI (t)  /  +10 €/t",
  "l(cpi10, 1)"  = "CPI (t–1)",
  "l(cpi10, 2)"  = "CPI (t–2)",
  "log_pop"      = "log(Population)",
  "log_gdp_pc"   = "log(GDP per capita)",
  "lgdppc_c"     = "Δ log GDPpc (ctr.)",
  "ind_c"        = "Δ Industry share (ctr.)",
  "ren_c"        = "Δ Renewable share (ctr.)"
)

# build the "Cumulative CPI (t..t+2)" extra row
row_main   <- cum_wald(m_main,   c("cpi10","l(cpi10, 1)","l(cpi10, 2)"))
row_robust <- cum_wald(m_robust, c("cpi10","l(cpi10, 1)","l(cpi10, 2)"))

add_rows_df <- tibble::tibble(
  term = "Cumulative CPI (t..t+2)",
  `Per-capita (main)`   = fmt_cell(row_main$est, row_main$se),
  `Totals (robustness)` = fmt_cell(row_robust$est, row_robust$se)
)

modelsummary(
  list("Per-capita (main)" = m_main,
       "Totals (robustness)" = m_robust),
  coef_map  = cm,
  gof_omit  = "IC|Log|Adj|F$|Std|^FE",
  statistic = "({std.error})",
  stars     = TRUE,
  add_rows  = add_rows_df,
  output    = "table_cpi_lags.tex",   # change to "markdown" or "html" if you prefer
  notes     = "Country & year FE in all columns. SE: Driscoll–Kraay with lag=3. Outcome is log points; CPI regressor is scaled so +1 = +10 €/t."
)

```

Local-projections (IRF) table — one coefficient per horizon

This shows the effect of ΔCPI (+10 €/t) at horizons h=0..3, keeping the other controls in the regression but only printing the ΔCPI coefficient.
```{r}
# 'mods' is your list: mods <- lapply(0:3, run_lp)
modelsummary(
  mods,
  coef_map  = c("dcpi" = "ΔCPI (+10 €/t)"),
  coef_omit = "^(?!dcpi$).*",  # keep only dcpi; regex needs PCRE (default)
  gof_omit  = "IC|Log|Adj|F$|Std|^FE",
  statistic = "({std.error})",
  stars     = TRUE,
  headers   = paste0("h=", 0:3),
  output    = "table_lp_irf.tex",
  notes     = "Local projections of y_{t+h} on ΔCPI_t (+10 €/t), with controls, country & year FE; SE: Driscoll–Kraay (lag=3)."
)

```


3) PDL (degree-2, lags 0..3): implied lag betas & cumulatives

Turns your m_pdl into a compact table of implied lag coefficients with 95% CIs, plus cumulative 0–2 and 0–3 rows.

```{r}
# rebuild implied lag betas & CIs
a <- coef(m_pdl)[c("pdl0","pdl1","pdl2")]
V <- vcov(m_pdl)[c("pdl0","pdl1","pdl2"), c("pdl0","pdl1","pdl2")]

k <- 0:3
b_hat <- as.numeric(a[1] + a[2]*k + a[3]*k^2)
J <- cbind(1, k, k^2)
se_b <- sqrt(diag(J %*% V %*% t(J)))
lo   <- b_hat - 1.96*se_b
hi   <- b_hat + 1.96*se_b

lag_df <- tibble::tibble(
  `Lag (years)` = paste0("t+", k),
  `Effect`      = sprintf("%.3f", b_hat),
  `SE`          = sprintf("(%.3f)", se_b),
  `95% CI`      = sprintf("[%.3f, %.3f]", lo, hi)
)

# cumulatives (0–2 and 0–3) via delta method
Sigma_b <- J %*% V %*% t(J)
L02 <- matrix(c(1,1,1,0), 1); L03 <- matrix(c(1,1,1,1), 1)

est02 <- as.numeric(L02 %*% b_hat); se02 <- sqrt(as.numeric(L02 %*% Sigma_b %*% t(L02)))
est03 <- as.numeric(L03 %*% b_hat); se03 <- sqrt(as.numeric(L03 %*% Sigma_b %*% t(L03)))

cum_df <- tibble::tibble(
  `Lag (years)` = c("Cumulative t..t+2", "Cumulative t..t+3"),
  `Effect`      = sprintf("%.3f", c(est02, est03)),
  `SE`          = sprintf("(%.3f)", c(se02, se03)),
  `95% CI`      = c(
    sprintf("[%.3f, %.3f]", est02 - 1.96*se02, est02 + 1.96*se02),
    sprintf("[%.3f, %.3f]", est03 - 1.96*se03, est03 + 1.96*se03)
  )
)

# bind and export
tbl_pdl <- dplyr::bind_rows(
  tibble::tibble(`Lag (years)` = "Panel A: Implied lag coefficients"), lag_df,
  tibble::tibble(`Lag (years)` = "Panel B: Cumulative effects"), cum_df
)

modelsummary::datasummary_df(
  tbl_pdl,
  output = "table_pdl_lags.tex",
  notes  = "PDL degree-2 over CPI lags 0..3. +1 in CPI regressor = +10 €/t. Country & year FE; SE: Driscoll–Kraay (lag=3)."
)

```


## RQ 5 
Phase IV (RQ5): Are effects stronger in Phase IV?
```{r}


# -----------------------
# 0) Prep: Phase IV flag
# -----------------------
as_ph <- analysis_sample %>%
  mutate(phaseIV = year >= 2021)  # TRUE for 2021–2023

# -----------------------
# 1) DL with Phase-IV interactions
# -----------------------
m_phase <- feols(
  log_co2_pc ~ l(cpi10, 0:2)*phaseIV + lgdppc_c + ind_c + ren_c | country + year,
  data = as_ph, panel.id = ~ country + year, vcov = vcov_DK(~year, lag = 3)
)

# -----------------------
# 2) Make cumulative rows (t..t+2) for baseline (pre-IV) and Phase IV = baseline + interaction
# -----------------------
cum_row_phase <- function(mod){
  nm  <- names(coef(mod)); V <- vcov(mod); b <- coef(mod)

  # Base terms (Phase II/III)
  base_terms <- nm[nm %in% c("cpi10","l(cpi10, 1)","l(cpi10, 2)")]
  # Interaction terms with Phase IV (names typically end with ':phaseIVTRUE')
  inter_terms <- nm[grepl("^cpi10:phaseIVTRUE$|^l\\(cpi10, 1\\):phaseIVTRUE$|^l\\(cpi10, 2\\):phaseIVTRUE$", nm)]

  vec_L <- function(terms){
    L <- rep(0, length(nm)); L[match(terms, nm)] <- 1; L
  }

  # Baseline cumulative
  L_base <- vec_L(base_terms)
  est_b  <- sum(b[match(base_terms, nm)])
  se_b   <- sqrt(drop(t(L_base) %*% V %*% L_base))

  # Phase IV cumulative = base + interactions
  both_terms <- c(base_terms, inter_terms)
  L_iv <- vec_L(both_terms)
  est_iv <- sum(b[match(both_terms, nm)])
  se_iv  <- sqrt(drop(t(L_iv) %*% V %*% L_iv))

  list(
    baseline = sprintf("%.3f\n(%.3f)", est_b, se_b),
    phaseIV  = sprintf("%.3f\n(%.3f)", est_iv, se_iv)
  )
}
rows_phase <- cum_row_phase(m_phase)

add_rows_phase <- tibble::tibble(
  term = c("Cumulative CPI (t..t+2), pre-Phase IV",
           "Cumulative CPI (t..t+2), Phase IV"),
  `Phase IV (per-capita)` = c(rows_phase$baseline, rows_phase$phaseIV)
)

# -----------------------
# 3) Publishable table
# -----------------------
cm_phase <- c(
  "cpi10"                 = "CPI (t)  /  +10 €/t",
  "l(cpi10, 1)"           = "CPI (t–1)",
  "l(cpi10, 2)"           = "CPI (t–2)",
  "cpi10:phaseIVTRUE"     = "× Phase IV",
  "l(cpi10, 1):phaseIVTRUE" = "× Phase IV (t–1)",
  "l(cpi10, 2):phaseIVTRUE" = "× Phase IV (t–2)",
  "lgdppc_c"              = "Δ log GDPpc (ctr.)",
  "ind_c"                 = "Δ Industry share (ctr.)",
  "ren_c"                 = "Δ Renewable share (ctr.)"
)

modelsummary(
  list("Phase IV (per-capita)" = m_phase),
  coef_map  = cm_phase,
  gof_omit  = "IC|Log|Adj|F$|Std|^FE",
  statistic = "({std.error})",
  stars     = TRUE,
  add_rows  = add_rows_phase,
  notes     = "Country & year FE. SE: Driscoll–Kraay (lag=3). +1 in CPI regressor = +10 €/t.",
  output    = "table_phaseIV.tex"   # change to "markdown" or "html" if you prefer
)

```

## RQ 2
Heterogeneity (RQ2): High-intensity & Low-renewables countries

Build pre-period groups (e.g., 2004–2006) and interact the CPI lags. Then compute subgroup-specific cumulative effects (0..2y).

```{r}
# -----------------------
# 0) Pre-period baseline groups
# -----------------------
base <- analysis_sample %>%
  filter(year %in% 2004:2006) %>%
  group_by(country) %>%
  summarise(
    base_int = mean(emis_int, na.rm=TRUE),
    base_ren = mean(renewable_share, na.rm=TRUE)
  ) %>%
  mutate(
    hi_int = base_int >= median(base_int, na.rm=TRUE),   # high emissions intensity
    lo_ren = base_ren <= median(base_ren, na.rm=TRUE)    # low renewable share
  )

as_het <- analysis_sample %>% left_join(base, by = "country")

# -----------------------
# 1) DL × High-intensity
# -----------------------
m_hi <- feols(
  log_co2_pc ~ l(cpi10, 0:2)*hi_int + lgdppc_c + ind_c + ren_c | country + year,
  data = as_het, panel.id = ~ country + year, vcov = vcov_DK(~year, lag = 3)
)

# -----------------------
# 2) DL × Low-renewables
# -----------------------
m_lr <- feols(
  log_co2_pc ~ l(cpi10, 0:2)*lo_ren + lgdppc_c + ind_c + ren_c | country + year,
  data = as_het, panel.id = ~ country + year, vcov = vcov_DK(~year, lag = 3)
)

# -----------------------
# 3) Cumulative rows helper (baseline group vs interacted group)
# -----------------------
cum_row_group <- function(mod, group_name){  # group_name = "hi_int" or "lo_ren"
  nm <- names(coef(mod)); V <- vcov(mod); b <- coef(mod)

  base_terms <- nm[nm %in% c("cpi10","l(cpi10, 1)","l(cpi10, 2)")]
  inter_terms <- nm[grepl(paste0(":", group_name, "TRUE$"), nm) &
                    grepl("^cpi10|^l\\(cpi10, 1\\)|^l\\(cpi10, 2\\)", nm)]

  vec_L <- function(terms){ L <- rep(0, length(nm)); L[match(terms, nm)] <- 1; L }

  # Baseline (group==FALSE)
  L_b <- vec_L(base_terms)
  est_b <- sum(b[match(base_terms, nm)])
  se_b  <- sqrt(drop(t(L_b) %*% V %*% L_b))

  # Group TRUE: base + interactions
  both <- c(base_terms, inter_terms)
  L_g <- vec_L(both)
  est_g <- sum(b[match(both, nm)])
  se_g  <- sqrt(drop(t(L_g) %*% V %*% L_g))

  list(
    baseline = sprintf("%.3f\n(%.3f)", est_b, se_b),
    group    = sprintf("%.3f\n(%.3f)", est_g, se_g)
  )
}

rows_hi <- cum_row_group(m_hi, "hi_int")
rows_lr <- cum_row_group(m_lr, "lo_ren")

add_rows_hi <- tibble::tibble(
  term = c("Cumulative CPI (t..t+2): low-intensity",
           "Cumulative CPI (t..t+2): high-intensity"),
  `By emissions intensity` = c(rows_hi$baseline, rows_hi$group)
)

add_rows_lr <- tibble::tibble(
  term = c("Cumulative CPI (t..t+2): high-renewables",
           "Cumulative CPI (t..t+2): low-renewables"),
  `By renewable share` = c(rows_lr$baseline, rows_lr$group)
)

# -----------------------
# 4) Publishable tables
# -----------------------
cm_het <- c(
  "cpi10"                   = "CPI (t)  /  +10 €/t",
  "l(cpi10, 1)"             = "CPI (t–1)",
  "l(cpi10, 2)"             = "CPI (t–2)",
  "cpi10:hi_intTRUE"        = "× High intensity",
  "l(cpi10, 1):hi_intTRUE"  = "× High intensity (t–1)",
  "l(cpi10, 2):hi_intTRUE"  = "× High intensity (t–2)",
  "cpi10:lo_renTRUE"        = "× Low renewables",
  "l(cpi10, 1):lo_renTRUE"  = "× Low renewables (t–1)",
  "l(cpi10, 2):lo_renTRUE"  = "× Low renewables (t–2)",
  "lgdppc_c"                = "Δ log GDPpc (ctr.)",
  "ind_c"                   = "Δ Industry share (ctr.)",
  "ren_c"                   = "Δ Renewable share (ctr.)"
)

modelsummary(
  list("By emissions intensity" = m_hi),
  coef_map  = cm_het,
  gof_omit  = "IC|Log|Adj|F$|Std|^FE",
  statistic = "({std.error})",
  stars     = TRUE,
  add_rows  = add_rows_hi,
  notes     = "Country & year FE. SE: Driscoll–Kraay (lag=3). +1 in CPI regressor = +10 €/t. Pre-period for grouping: 2004–2006.",
  output    = "table_heterogeneity_intensity.tex"
)

modelsummary(
  list("By renewable share" = m_lr),
  coef_map  = cm_het,
  gof_omit  = "IC|Log|Adj|F$|Std|^FE",
  statistic = "({std.error})",
  stars     = TRUE,
  add_rows  = add_rows_lr,
  notes     = "Country & year FE. SE: Driscoll–Kraay (lag=3). +1 in CPI regressor = +10 €/t. Pre-period for grouping: 2004–2006.",
  output    = "table_heterogeneity_renewables.tex"
)


```
### RQ summaries
Phase IV (RQ5):
“Interacting CPI lags with Phase IV (2021–2023) shows that the cumulative 0–2-year effect becomes more negative in Phase IV than in earlier phases (Table ‘Phase IV’). While contemporaneous effects remain indistinguishable from zero, the one- and two-year responses are larger in magnitude under tighter caps and higher auctioning.”

Heterogeneity (RQ2):
“Cumulative 0–2-year effects are more negative in high-intensity and low-renewables countries (Table ‘By emissions intensity’ / ‘By renewable share’), consistent with greater abatement opportunities and stronger responses where carbon intensity is initially high or clean energy penetration is initially low.”

# Additional analysis

Local Projections × Phase IV (RQ5)

What you’ll get: one table with horizons h = 0,1,2 showing:

the baseline ΔCPI effect (pre-Phase IV),

the differential ΔCPI × Phase IV,

and an extra row “Effect in Phase IV (sum)” = baseline + differential with its SE (delta method).


```{r}
# 0) Clean keys & CPI change ( +1 == +10 €/t )
analysis_sample$year <- as.integer(analysis_sample$year)
stopifnot(!any(duplicated(analysis_sample[c("country","year")])))

as_lp <- analysis_sample %>%
  arrange(country, year) %>%
  group_by(country) %>%
  mutate(dcpi = cpi10 - lag(cpi10)) %>%
  ungroup() %>%
  mutate(phaseIV = year >= 2021)

# 1) LP runner with Phase IV interaction
run_lp_phase <- function(h){
  dat_h <- as_lp %>%
    group_by(country) %>%
    mutate(y_h = dplyr::lead(log_co2_pc, h)) %>%
    ungroup()
  feols(y_h ~ dcpi * phaseIV + lgdppc_c + ind_c + ren_c | country + year,
        data = dat_h,
        vcov = vcov_DK(~ year, lag = 3))
}

mods_ph <- lapply(0:2, run_lp_phase)
names(mods_ph) <- paste0("h=", 0:2)

# 2) Helper: baseline ΔCPI and Phase IV "sum" (ΔCPI + ΔCPI×PhaseIV) with SE (delta)
effect_sum <- function(mod, base_pat="^dcpi$", inter_pat="^dcpi:.*phaseIV.*$"){
  nm <- names(coef(mod)); b <- coef(mod); V <- vcov(mod)
  base <- nm[grepl(base_pat, nm)]
  inter <- nm[grepl(inter_pat, nm)]
  # baseline
  est_b <- sum(b[base]); se_b <- sqrt(drop(t(1) %*% V[base, base, drop=FALSE] %*% 1))
  # sum (baseline + interaction)
  idx <- c(base, inter)
  L <- matrix(1, nrow=1, ncol=length(idx))
  est_s <- sum(b[idx]); se_s <- sqrt(drop(L %*% V[idx, idx, drop=FALSE] %*% t(L)))
  list(baseline = sprintf("%.3f\n(%.3f)", est_b, se_b),
       phaseIV  = sprintf("%.3f\n(%.3f)", est_s, se_s))
}

rows <- lapply(mods_ph, effect_sum)

add_rows_phase <- tibble::tibble(
  term = c("Effect in Phase IV (ΔCPI + ΔCPI×Phase IV)"),
  !!!setNames(lapply(rows, `[[`, "phaseIV"), names(mods_ph))
)

# 3) Publishable table (only print ΔCPI and interaction in the body; sum goes in add_rows)
cm_phase <- c("dcpi" = "ΔCPI (+10 €/t)",
              "dcpi:phaseIVTRUE" = "ΔCPI × Phase IV")

modelsummary(
  mods_ph,
  coef_map  = cm_phase,
  coef_omit = "^(?!dcpi$|dcpi:phaseIVTRUE$).*",  # keep only these two
  gof_omit  = "IC|Log|Adj|F$|Std|^FE",
  statistic = "({std.error})",
  stars     = TRUE,
  add_rows  = add_rows_phase,
  notes     = "Local projections of y_{t+h} on ΔCPI_t (+10 €/t), with country & year FE and controls; SE: Driscoll–Kraay (lag=3).",
  output    = "table_lp_phaseIV.tex"
)

```

