---
title: "wp_carbon"
author: "sara"
date: "2025-10-04"
output: html_document
---

# Abstract
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setwd("~/Library/CloudStorage/OneDrive-umontpellier.fr/Project 2/carbon analysis")
knitr::opts_knit$set(root.dir = "/Users/saramustafazade/Library/CloudStorage/OneDrive-umontpellier.fr/Project 2/carbon analysis")

```

# Data loading and cleaning

```{r}
# Packages
library(skimr)
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(stringr)
library(lubridate)
library(car)
# to pull variables
library(eurostat)
library(countrycode)
library(WDI)
# for regression
library(fixest)
library(broom)
library(fixest)
library(modelsummary)
library(glue)
library(plm)
library(sandwich); library(lmtest)
library(tibble)
library(kableExtra)

# Load the data

# read the sheet called "final_sheet"
df <- read_excel("carbon_index_2025.xlsx", sheet = "final_sheet")
co2<- read_excel("IEA_EDGAR_CO2_1970_2023.xlsx", sheet = "TOTALS BY COUNTRY")

# check the data
#head(df)
#head(oecd)
```

Filetiring EU countries from our dataset
```{r}

# vector of EU countries
eu_countries <- c(
  "Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus", "Czech Republic",
  "Denmark", "Estonia", "Finland", "France", "Germany", "Greece", "Hungary",
  "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg", "Malta",
  "Netherlands", "Poland", "Portugal", "Romania", "Slovakia", "Slovenia", 
  "Spain", "Sweden"
)
# filter only EU countries
df_eu <- df %>% 
  filter(ds_country %in% eu_countries)  

co2_eu <- co2 %>%
  filter(Name %in% eu_countries)



# check
#head(df_eu)
#head(co2_eu)
```

```{r}
# create a summary table showing the earliest start year for each mechanism in each country
summary_tbl <- df_eu %>% 
  group_by(ds_country, implemented_mechanisms) %>% 
  summarize(start_year = min(start_year, na.rm = TRUE), .groups = "drop" )

#view the summary table
summary_tbl
```

```{r}
# Preparing CO2 emission data for merging
co2_eu_filtered <- co2_eu %>%
  select(IPCC_annex, C_group_IM24_sh, Country_code_A3, Name, Substance,
         Y_2000:Y_2023)

# Reshaping CO2 data from wide to long format

co2_eu_long <- co2_eu %>%
  pivot_longer(cols = starts_with("Y_"),
               names_to = "year",
               values_to = "co2") %>%
  mutate(year = as.integer(sub("Y_", "", year))) %>%
  filter(year >= 2000 & year <= 2023)

# CO2 table: standardize to the same key ---
co2_key <- co2_eu_long %>%
  filter(Substance == "CO2") %>%
  rename(iso3 = Country_code_A3) %>%
  select(iso3, year, co2)   # keep only what's needed

```

```{r df_eu to long}
# ==============================================================================
# ## Step 1: Pivot the Carbon Pricing Index Data to a Long Format
# ==============================================================================
# We select the country identifiers and all columns that start with "carbon_pricing_index_"

cpi_long <- df_eu %>%
  select(ds_country, iso3, starts_with("carbon_pricing_index_")) %>%
  pivot_longer(
    cols = -c(ds_country, iso3),
    names_to = "year",
    values_to = "carbon_pricing_index"
  ) %>%
  mutate(year = str_remove(year, "carbon_pricing_index_"))%>%
  mutate(year = as.integer(year))
# ==============================================================================
# ## Step 3: Join the Two Long Datasets into a Final Panel
# ==============================================================================

# We join by country and year to bring everything together.
# We also convert 'year' to a numeric type for easier analysis.
panel_df <- co2_key %>%
  full_join(cpi_long, by = c( "iso3", "year")) %>%
  mutate(year = as.numeric(year)) %>%
  arrange(iso3, year)

# --- View the result ---
cat("Data has been successfully transformed into a long panel format:\n")
head(panel_df)


```

### Pulling control variables

1) Pull from WDI

```{r}

# Step 1: correct indicator codes
wdi_vars <- c("NY.GDP.MKTP.KD",   # GDP (constant 2015 US$)
              "SP.POP.TOTL",      # Population
              "NV.IND.TOTL.ZS")   # Industry share of GDP (%)

# Step 2: fetch from World Bank
wdi_raw <- WDI(
  country = "all",
  indicator = wdi_vars,
  start = 2000,
  end = 2025
)
# Step 3: Clean and rename (Corrected version)
wdi_clean <- wdi_raw %>%
  rename(
    iso3 = iso3c,  # Use the existing iso3c column as the identifier
    year = year,
    gdp = NY.GDP.MKTP.KD,
    population = SP.POP.TOTL,
    industry_share = NV.IND.TOTL.ZS
  ) %>%
  select(country, iso3, year, gdp, population, industry_share) # Keep the original country name and the new iso3

# Define EU countries in ISO3 codes
eu_iso3 <- c("AUT","BEL","BGR","HRV","CYP","CZE","DNK","EST","FIN","FRA","DEU",
             "GRC","HUN","IRL","ITA","LVA","LTU","LUX","MLT","NLD","POL","PRT",
             "ROU","SVK","SVN","ESP","SWE")

# Keep only EU countries using the clean 'iso3' column
wdi_eu <- wdi_clean %>%
  filter(iso3 %in% eu_iso3)

# Quick check to confirm
head(wdi_eu)


#############

# 2) Merge with your dataset
merged_controls <- panel_df %>%
  left_join(wdi_eu, by = c("iso3", "year"))


# 3) Create transformed controls
merged_controls <- merged_controls %>%
  mutate(
    log_gdp = log(gdp),
    log_pop = log(population),
    gdp_per_capita = gdp / population,
    log_gdp_pc = log(gdp_per_capita)
  )

```
Pulling renewable energy share from Eurostat.
nrg_ind_ren, is one of Eurostat's main energy indicators.
```{r}

# EU-27 (Greece is "EL" in Eurostat)
eu_geo <- c("AT","BE","BG","HR","CY","CZ","DK","EE","FI","FR","DE","EL","HU",
            "IE","IT","LV","LT","LU","MT","NL","PL","PT","RO","SK","SI","ES","SE")

# Pull raw (cache speeds things up), detect the time col (time or TIME_PERIOD)
renewables_raw <- get_eurostat("nrg_ind_ren", cache = TRUE)  # time can be Date or TIME_PERIOD
time_col <- intersect(c("time","TIME_PERIOD"), names(renewables_raw))[1]

# Collapse to one obs per country–year, keep unit == % (PC)
renewables_eu <- renewables_raw %>%
  filter(geo %in% eu_geo, unit == "PC") %>%
  group_by(geo, !!sym(time_col)) %>%
  summarise(renewable_share_pc = mean(values, na.rm = TRUE), .groups = "drop") %>%
  rename(year_raw = !!sym(time_col)) %>%
  mutate(
    year    = as.integer(format(as.Date(year_raw), "%Y")),
    year    = ifelse(is.na(year), as.integer(year_raw), year),
    # Create the 'iso3' column directly with the correct name
    iso3 = countrycode(geo, "eurostat", "iso3c") 
  ) %>%
  # Select only the columns needed, using the 'iso3' name
  select(iso3, year, renewable_share_pc) %>% 
  arrange(iso3, year)

# (The rest of the script for padding and merging remains the same)

# The final merge will now work perfectly
merged_final <- merged_controls %>%
  left_join(renewables_eu, by = c("iso3", "year")) # Merging by 'iso3'
# Safety checks
stopifnot(!anyDuplicated(renewables_eu[c("iso3","year")]))

# Optional: pad to full 2000–2025 panel
renewables_eu <- renewables_eu %>%
  complete(iso3, year = 2000:2025) %>%
  mutate(renewable_share = renewable_share_pc / 100)  # fraction 0–1 (keep both if you like)

# Merge with your main panel (iso3, year)
merged_final <- merged_controls %>%
  left_join(renewables_eu, by = c("iso3","year"))

# Sanity Checks (if necessery uncomment)

# # Post-merge sanity: duplicates? missingness?
# stopifnot(!anyDuplicated(merged_final[c("iso3","year")]))
# 
# # coverage summary
# renewables_eu %>%
#   group_by(iso3) %>%
#   summarise(n_years = n_distinct(year),
#             min_year = min(year), max_year = max(year), .groups = "drop") %>%
#   arrange(iso3)
# 
# nrow(renewables_eu)        # expect ~27 * 20 = 540 (2004–2023)
# length(unique(renewables_eu$iso3))  # expect 27
# length(unique(renewables_eu$year))     # expect 20


```

```{r}

#  Create ONLY the missing outcome variants (keep your existing log_gdp_pc)
merged_final <- merged_final %>%
  mutate(
    # safe logs (NA if non-positive)
    log_co2     = ifelse(co2 > 0, log(co2), NA_real_),
    co2_pc      = co2 / population,
    log_co2_pc  = ifelse(co2_pc > 0, log(co2_pc), NA_real_),
    emis_int    = co2 / gdp,              # emissions intensity
    # optional: fraction 0–1 if you kept renewable_share_pc earlier
    renewable_share = ifelse(!is.na(renewable_share), renewable_share,
                             ifelse(exists("renewable_share_pc"),
                                    renewable_share_pc/100, NA_real_))
  )

# Define your main estimation sample (overlap years; drop NAs in key vars)
analysis_sample <- merged_final %>%
  filter(year >= 2004, year <= 2023) %>%     # renewables start in 2004
  tidyr::drop_na(log_co2, carbon_pricing_index, log_gdp_pc, industry_share, renewable_share)

# Scales & centering (for interpretability & interactions)
analysis_sample <- analysis_sample %>%
  mutate(
    cpi10   = carbon_pricing_index / 10,            # 10-pt change
    lcpi    = log1p(carbon_pricing_index),          # optional robustness
    ren_c   = scale(renewable_share, scale = FALSE),# mean-centered
    lgdppc_c= scale(log_gdp_pc, scale = FALSE),
    ind_c   = scale(industry_share, scale = FALSE),
    t       = year                                   # for trends
  )

```

# Analysis

## RQ1

```{r}
m_rq1_pc <- feols(
log_co2_pc ~ cpi10 + lgdppc_c + ind_c + ren_c | country + year,
data= analysis_sample,
panel.id = ~ country + year,
vcov= ~ country + year # two-way clustered SE
#vcov = vcovDK(time = ~ year, kernel = "Bartlett", bandwidth = 3)
)

# ----- RQ1 robustness: Totals (add log_pop), same within-controls -----
m_rq1_tot <- feols(
log_co2 ~ cpi10 + log_pop + lgdppc_c + ind_c + ren_c | country + year,
data = analysis_sample,
panel.id = ~ country + year,
vcov = ~ country + year,
# vcov = vcovDK(time = ~ year, kernel = "Bartlett", bandwidth = 3)
)

etable(m_rq1_pc, m_rq1_tot, se.below = TRUE,
       headers = c("Per-capita (main)", "Totals (robustness)"))

```
```{r}

models_rq1 <- list(
  "Per-capita (main)"      = m_rq1_pc,
  "Per-capita + Trends"    = m_rq1_pc_trend,
  "Totals (robustness)"    = m_rq1_tot,
  "No Renewables"          = m_pc_noren
)

msummary(
  models_rq1,
  fmt = 3,
  stars = c('*'=.05,'**'=.01,'***'=.001),
  # coef_map = c(
  #   "cpi10"     = "CPI (+10 €/t)",
  #   "lgdppc_c"  = "log GDP pc (within)",
  #   "ind_c"     = "Industry share",
  #   "ren_c"     = "Renewable share",
  #   "log_pop"   = "log Population"
  # ),
  # gof_map = c("n","r.squared.within"),
  notes = "Country and year fixed effects in all models. Standard errors clustered two-ways by country and year. Significance: * p < 0.05, ** p < 0.01, *** p < 0.001.",
  output = "latex"
)

```
The coefficients on CPI are negative in all models but not statistically significant, implying no immediate reduction in emissions from higher carbon pricing.

Given the scaling (CPI = +10 €/t), even the largest point estimate (−0.007) implies only about 0.7 % lower per-capita emissions per +10 €/t, economically small and statistically uncertain.

Renewable share is consistently large and negative (−1.1 to −1.7, p < 0.01), confirming that clean-energy expansion explains most of the contemporaneous decarbonization.

Adding country-specific trends absorbs some of this variation but does not change the overall conclusion.

```{r app: country trend}

m_rq1_pc_trend <- feols(
  log_co2_pc ~ cpi10 + lgdppc_c + ind_c + ren_c + i(country, year) | country + year,
  data = analysis_sample, vcov = ~ country + year
)
etable(m_rq1_pc, m_rq1_pc_trend, se.below = TRUE,
       headers = c("No trend", "Country-specific trends"))

```
```{r app: renewables}
m_pc_noren <- feols(
  log_co2_pc ~ cpi10 + lgdppc_c + ind_c | country + year,
  data = analysis_sample, vcov = ~ country + year
)
etable(m_rq1_pc, m_pc_noren, headers = c("With renewables","No renewables"))

```
```{r app: high price}
analysis_sample <- analysis_sample %>% mutate(high_price = as.integer(carbon_pricing_index >= 30))
m_rq1_hp <- feols(
  log_co2_pc ~ cpi10*high_price + lgdppc_c + ind_c + ren_c | country + year,
  data = analysis_sample, vcov = ~ country + year
)
# Marginal effect at high prices:
etable(m_rq1_pc, m_rq1_hp, se.below = TRUE,
       headers = c("All prices","With high-price interaction"))

```
## RQ 2

```{r}
# --- RQ2: Dynamics of CPI (immediate vs lagged) ---

# Short DL (0–2): contemporaneous + 1–2 year lags
m_rq2_dl02 <- feols(
  log_co2_pc ~ l(cpi10, 0:2) + lgdppc_c + ind_c + ren_c | country + year,
  data      = analysis_sample,
  panel.id  = ~ country + year,         # <-- REQUIRED for l()
  vcov      = ~ country + year          # or: vcovDK(time = ~year, bandwidth = 3)
)

# Placebo leads to check pre-trends (should be ~0 if no anticipation)
m_rq2_leads_lags <- feols(
  log_co2_pc ~ f(cpi10, 1:2) + l(cpi10, 0:2) + lgdppc_c + ind_c + ren_c | country + year,
  data      = analysis_sample,
  panel.id  = ~ country + year,
  vcov      = ~ country + year
)

# Cumulative (sum of 0–2 lags) Wald test
#wald(m_rq2_dl02, "l(cpi10,0) + l(cpi10,1) + l(cpi10,2) = 0")

# Nicely formatted table
etable(m_rq2_dl02, m_rq2_leads_lags, se.below = TRUE,
       headers = c("DL(0–2)", "Leads(1–2) + DL(0–2)"))

```

```{r}
models_rq2 <- list(
  "Per-capita (main)"      = m_rq2_dl02,
  "Per-capita + Trends"    = m_rq2_leads_lags
)

msummary(
  models_rq2,
  fmt = 3,
  stars = c('*'=.05,'**'=.01,'***'=.001),
  # #coef_map = c(
  #   "cpi10"     = "CPI (+10 €/t)",
  #   "lgdppc_c"  = "log GDP pc (within)",
  #   "ind_c"     = "Industry share",
  #   "ren_c"     = "Renewable share",
  #   "log_pop"   = "log Population"
  # ),
  # gof_map = c("n","r.squared.within"),
  notes = "Dynamics of CPI (immediate vs lagged). Country and year fixed effects in all models. Standard errors clustered two-ways by country and year. Significance: * p < 0.05, ** p < 0.01, *** p < 0.001.",
  output = "html"
)

```

```{r wald rq2}
# Robust cumulative-effect test for lags 0:k (k >= 0)
cum_wald_auto <- function(mod, var = "cpi10", lags = 0:2) {
  nm <- names(coef(mod))
  want <- character()

  for (k in lags) {
    if (k == 0) {
      # contemporaneous is just the bare var name
      cand <- var
    } else {
      # possible printed names: "l(cpi10,1)" or "l(cpi10, 1)"
      cand <- c(sprintf("l(%s,%d)", var, k),
                sprintf("l(%s, %d)", var, k))
    }
    hit <- cand[cand %in% nm]
    if (length(hit) == 0) {
      stop("Could not find coefficient for lag ", k,
           ". Tried: ", paste(cand, collapse = " | "),
           "\nAvailable names include e.g.: ", paste(head(nm, 10), collapse = ", "), " ...")
    }
    want <- c(want, hit[1])
  }

  b <- coef(mod)[want]
  V <- vcov(mod)[want, want, drop = FALSE]
  L <- matrix(1, nrow = 1, ncol = length(want))

  est <- as.numeric(L %*% b)
  se  <- sqrt(as.numeric(L %*% V %*% t(L)))
  z   <- est / se
  p   <- 2 * pnorm(abs(z), lower.tail = FALSE)
  ci  <- est + c(-1, 1) * 1.96 * se
  pct <- 100 * (exp(est) - 1)     # percent interpretation for log DV

  list(est = est, se = se, z = z, p = p, ci = ci, pct = pct, terms = want)
}

# ---- Apply to your RQ2 models ----
#rq2_sum     <- cum_wald_auto(m_rq2_dl02,        var = "cpi10", lags = 0:2)
#rq2_sum_lds <- cum_wald_auto(m_rq2_leads_lags,  var = "cpi10", lags = 0:2)

#rq2_sum
#rq2_sum_lds
# see below for table of these results

```
```{r rq2 table}
# cumulative test (from your helper)
cw_dl   <- cum_wald_auto(m_rq2_dl02, var = "cpi10", lags = 0:2)
cw_lead <- cum_wald_auto(m_rq2_leads_lags, var = "cpi10", lags = 0:2)

# create a small data frame to add as a custom note line
cum_line <- data.frame(
  Variable = "Cumulative CPI (0–2y)",
  `DL(0–2)`           = sprintf("%.3f (%.3f)", cw_dl$est, cw_dl$se),
  `Leads+DL(0–2)`     = sprintf("%.3f (%.3f)", cw_lead$est, cw_lead$se)
)

# show main etable and append manually below
etable(m_rq2_dl02, m_rq2_leads_lags,
       headers = c("DL(0–2)", "Leads(1–2)+DL(0–2)"),
       se.below = TRUE)

print(cum_line, row.names = FALSE)


```


RQ2 (timing). Distributed-lag estimates over t,t−1,t−2 indicate that the contemporaneous CPI effect is close to zero, while the t−1 coefficient is negative and statistically significant (−0.015, p<0.05); the t−2 coefficient is small and imprecise. Placebo leads are insignificant, suggesting no pre-trends. The cumulative 0–2-year effect of a +10 €/t increase is about −1.0% (DL(0–2): −0.0103, SE 0.0205, p=0.615; Leads+DL: −0.0096, SE 0.0197, p=0.628). These patterns imply that most of the policy impact materializes with a one-year delay, consistent with investment and adoption cycles, while the multi-year average over two years is modest and imprecisely estimated.

```{r}
# Coefficient plot focusing on CPI terms
coefplot(
  m_rq2_leads_lags,
  keep = c("f\\(cpi10,1\\)","f\\(cpi10,2\\)","cpi10","l\\(cpi10,1\\)","l\\(cpi10,2\\)"),
  main = "Dynamic effects of CPI (+10 €/t) on log CO2 per capita",
  xlab = "Semi-elasticity",
  dict = c(
    "f(cpi10,1)" = "Lead 1",
    "f(cpi10,2)" = "Lead 2",
    "cpi10"      = "t",
    "l(cpi10,1)" = "t−1",
    "l(cpi10,2)" = "t−2"
  )
)


```
## RQ 3 

```{r}
# --- Construct HighCarbon indicator (pre-2004–2006 baseline) ---
baseline_intensity <- analysis_sample %>%
  filter(year %in% 2004:2006) %>%
  group_by(country) %>%
  summarise(emis_int_base = mean(emis_int, na.rm = TRUE), .groups = "drop") %>%
  mutate(HighCarbon = ifelse(emis_int_base > median(emis_int_base, na.rm = TRUE), 1, 0))

df_rq3 <- analysis_sample %>%
  left_join(baseline_intensity, by = "country")

# --- Heterogeneity regression ---
m_rq3 <- feols(
  log_co2_pc ~ cpi10 * HighCarbon + lgdppc_c + ind_c + ren_c | country + year,
  data = df_rq3, vcov = ~ country + year
)

etable(m_rq3, se.below = TRUE, headers = "H3: High vs. Low Carbon Intensity")

```

```{r not working}
# Effects and tests
# High group total effect = cpi10 + cpi10:HighCarbon
wald(m_rq3, "cpi10 + cpi10:HighCarbon = 0")   # high group
wald(m_rq3, "cpi10 = 0")                      # low group
wald(m_rq3, "cpi10:HighCarbon = 0")           # difference

```
Test dynamic heterogeneity (this is the one likely to confirm H3)
```{r}
# Dynamic heterogeneity: interact lags with HighCarbon
m_rq3_dyn <- feols(
  log_co2_pc ~ cpi10*HighCarbon + l(cpi10,1)*HighCarbon + l(cpi10,2)*HighCarbon +
               lgdppc_c + ind_c + ren_c | country + year,
  data = df_rq3,
  panel.id = ~ country + year,
  vcov = ~ country + year
)

# Cumulative (0–2y) effects by group:
# low group: sum of cpi10 + l(cpi10,1) + l(cpi10,2)
wald(m_rq3_dyn, "cpi10 + l(cpi10,1) + l(cpi10,2) = 0")

# high group: add the three interactions
wald(m_rq3_dyn, paste(
  "cpi10 + l(cpi10,1) + l(cpi10,2) +",
  "cpi10:HighCarbon + l(cpi10,1):HighCarbon + l(cpi10,2):HighCarbon = 0"
))

# difference in cumulative effects (High − Low)
wald(m_rq3_dyn, paste(
  "cpi10:HighCarbon + l(cpi10,1):HighCarbon + l(cpi10,2):HighCarbon = 0"
))
etable(m_rq3_dyn, se.below = TRUE,
       headers = "H3 Dynamic: High vs. Low Carbon Intensity")
```
Adding a continuous heterogeneity (more power, cleaner)
```{r just for check}
# Standardize baseline intensity (z-score)
df_rq3 <- df_rq3 |>
  mutate(z_intensity = as.numeric(scale(emis_int_base)))

m_rq3_cont <- feols(
  log_co2_pc ~ cpi10*z_intensity + lgdppc_c + ind_c + ren_c | country + year,
  data = df_rq3, vcov = ~ country + year
)

# Slope at −1 SD, mean, +1 SD:
sd1 <- 1
wald(m_rq3_cont, "cpi10 + (-1)*cpi10:z_intensity = 0")  # low baseline
wald(m_rq3_cont, "cpi10 = 0")                           # mean baseline
wald(m_rq3_cont, "cpi10 + ( 1)*cpi10:z_intensity = 0")  # high baseline

```

```{r}
# Find one term among multiple acceptable name variants
.find_one <- function(nm, candidates) {
  hit <- candidates[candidates %in% nm]
  if (length(hit) == 0) return(NA_character_)
  hit[1]
}

# Build name-robust list of terms for dynamic het. (0–2y)
get_terms_dynhet <- function(nm, var = "cpi10", lags = 0:2, int = "HighCarbon") {
  out_low  <- character(0)
  out_high <- character(0)

  for (k in lags) {
    if (k == 0) {
      # base contemporaneous
      base <- .find_one(nm, var)
      # interaction can be in either order, with/without spaces
      int_cands <- c(
        sprintf("%s:%s", var, int),
        sprintf("%s:%s", int, var)
      )
      inter <- .find_one(nm, int_cands)
    } else {
      # lag name variants with/without spaces
      lag_cands <- c(
        sprintf("l(%s,%d)", var, k),
        sprintf("l(%s, %d)", var, k)
      )
      base <- .find_one(nm, lag_cands)

      int_cands <- c(
        sprintf("l(%s,%d):%s", var, k, int),
        sprintf("l(%s, %d):%s", var, k, int),
        sprintf("%s:l(%s,%d)", int, var, k),
        sprintf("%s:l(%s, %d)", int, var, k)
      )
      inter <- .find_one(nm, int_cands)
    }
    out_low  <- c(out_low,  base)
    out_high <- c(out_high, inter)
  }

  list(low = out_low, high = out_high)
}

# Generic linear-combo (sum) with informative errors
lincombo_sum <- function(mod, terms) {
  nm <- names(coef(mod))
  if (anyNA(terms)) {
    miss_k <- which(is.na(terms))
    stop("Some requested terms were not found in the model:\n",
         paste(sprintf("  - component %d is missing", miss_k), collapse = "\n"),
         "\nInspect names(coef(model)) to see available terms.")
  }
  b <- coef(mod)[terms]
  V <- vcov(mod)[terms, terms, drop = FALSE]
  est <- sum(b)
  se  <- sqrt(sum(V))             # because all weights are 1
  z   <- est/se
  p   <- 2*pnorm(abs(z), lower.tail = FALSE)
  ci  <- est + c(-1,1)*1.96*se
  list(est=est, se=se, z=z, p=p, ci=ci, terms=terms)
}


```

```{r}
nm <- names(coef(m_rq3_dyn))
tt <- get_terms_dynhet(nm, var = "cpi10", lags = 0:2, int = "HighCarbon")

# (A) Low-intensity cumulative (0–2y): cpi10 + l1 + l2
res_low  <- lincombo_sum(m_rq3_dyn, tt$low)

# (B) High-intensity cumulative (0–2y): add the three interactions
res_high <- lincombo_sum(m_rq3_dyn, c(tt$low, tt$high))

# (C) Difference in cumulative effects (High − Low): just the three interactions
res_diff <- lincombo_sum(m_rq3_dyn, tt$high)

res_low
res_high
res_diff

```
```{r}
# ---- store your cumulative results ----
rq3_tab <- tribble(
  ~Group,         ~Estimate,    ~`Std. Error`, ~z,        ~`p-value`, ~`95% CI Lower`, ~`95% CI Upper`, ~`% Δ CO2 (cum.)`,
  "Low-intensity", -0.0128,      0.0223,       -0.573,    0.567,      -0.0565,         0.0310,          -1.28,
  "High-intensity",-0.0060,      0.0209,       -0.288,    0.773,      -0.0471,         0.0350,          -0.60,
  "Difference (H–L)", 0.0068,    0.0214,        0.315,    0.753,      -0.0353,         0.0488,           0.68
)

# ---- knit to LaTeX ----
kable(
  rq3_tab, "html", booktabs = TRUE, digits = 4,
  caption = "RQ3: Cumulative 0--2-year effects of Carbon Pricing Index (CPI) by baseline carbon intensity. 
Standard errors clustered two ways (country and year). CPI scaled by +10 €/t."
)


```
Does RQ3 hold with CO2 totals?

```{r}
# Assumes: df_rq3 has HighCarbon and (optionally) z_intensity from before
# If not, rebuild HighCarbon exactly as you did for per-capita.

# --- Dynamic heterogeneity with totals (log CO2) ---
m_rq3_dyn_tot <- feols(
  log_co2 ~ cpi10*HighCarbon + l(cpi10,1)*HighCarbon + l(cpi10,2)*HighCarbon +
            log_pop + lgdppc_c + ind_c + ren_c | country + year,
  data     = df_rq3,
  panel.id = ~ country + year,
  vcov     = ~ country + year
)

# --- Helper to grab correct coefficient names and sum them (as before) ---
.find_one <- function(nm, candidates){ hit <- candidates[candidates %in% nm]; if(!length(hit)) NA_character_ else hit[1] }
get_terms_dynhet <- function(nm, var="cpi10", lags=0:2, int="HighCarbon"){
  low <- high <- character(0)
  for(k in lags){
    if(k==0){
      base <- .find_one(nm, var)
      inter <- .find_one(nm, c(sprintf("%s:%s",var,int), sprintf("%s:%s",int,var)))
    } else {
      lag_cands <- c(sprintf("l(%s,%d)",var,k), sprintf("l(%s, %d)",var,k))
      base <- .find_one(nm, lag_cands)
      int_cands <- c(sprintf("l(%s,%d):%s",var,k,int), sprintf("l(%s, %d):%s",var,k,int),
                     sprintf("%s:l(%s,%d)",int,var,k), sprintf("%s:l(%s, %d)",int,var,k))
      inter <- .find_one(nm, int_cands)
    }
    low  <- c(low,  base)
    high <- c(high, inter)
  }
  list(low=low, high=high)
}
lincombo_sum <- function(mod, terms){
  nm <- names(coef(mod))
  if(anyNA(terms)) stop("Missing terms: ", paste(which(is.na(terms)), collapse=", "),
                        "\nCheck names(coef(model)).")
  b <- coef(mod)[terms]
  V <- vcov(mod)[terms, terms, drop=FALSE]
  est <- sum(b)
  se  <- sqrt(sum(V))     # all weights 1
  z   <- est/se
  p   <- 2*pnorm(abs(z), lower.tail = FALSE)
  ci  <- est + c(-1,1)*1.96*se
  list(est=est, se=se, z=z, p=p, ci=ci, terms=terms)
}

# --- Cumulative (0–2y): Low, High, Difference for totals ---
nm_tot <- names(coef(m_rq3_dyn_tot))
tt_tot <- get_terms_dynhet(nm_tot, var="cpi10", lags=0:2, int="HighCarbon")

res_low_tot  <- lincombo_sum(m_rq3_dyn_tot, tt_tot$low)
res_high_tot <- lincombo_sum(m_rq3_dyn_tot, c(tt_tot$low, tt_tot$high))
res_diff_tot <- lincombo_sum(m_rq3_dyn_tot, tt_tot$high)

res_low_tot
res_high_tot
res_diff_tot

# Optional: put these into an etable GOF block
etable(
  m_rq3_dyn_tot,
  se.below = TRUE,
  headers = "RQ3 (Totals): Dynamic heterogeneity (0–2y)"
  
)
gof = list(
    "Cumulative (Low, 0–2y)"  = sprintf("%.4f (%.4f), p=%.3f", res_low_tot$est,  res_low_tot$se,  res_low_tot$p),
    "Cumulative (High, 0–2y)" = sprintf("%.4f (%.4f), p=%.3f", res_high_tot$est, res_high_tot$se, res_high_tot$p),
    "Diff High−Low (0–2y)"    = sprintf("%.4f (%.4f), p=%.3f", res_diff_tot$est, res_diff_tot$se, res_diff_tot$p)
  )

```
```{r}

# --------------------------
# 0) (Optional) DK vcov (Driscoll–Kraay)
# --------------------------
# Use this if you want DK SEs instead of two-way cluster.
vDK <- function(x) vcovSCC(x, time = ~ year, kernel = "Bartlett", maxlag = 3)

# --------------------------
# 1) Helpers to robustly find names & compute linear combos
# --------------------------
.find_one <- function(nm, candidates){
  hit <- candidates[candidates %in% nm]
  if (!length(hit)) NA_character_ else hit[1]
}

# Collect coefficient names for dynamic heterogeneity: base + lags, and their interactions
get_terms_dynhet <- function(nm, var = "cpi10", lags = 0:2, int = "HighCarbon"){
  low <- high <- character(0)
  for (k in lags) {
    if (k == 0) {
      base   <- .find_one(nm, var)
      inter  <- .find_one(nm, c(sprintf("%s:%s", var, int), sprintf("%s:%s", int, var)))
    } else {
      # handle both l(cpi10,1) and l(cpi10, 1) (with a space)
      lag_cands <- c(sprintf("l(%s,%d)", var, k), sprintf("l(%s, %d)", var, k))
      base      <- .find_one(nm, lag_cands)

      int_cands <- c(sprintf("l(%s,%d):%s", var, k, int),
                     sprintf("l(%s, %d):%s", var, k, int),
                     sprintf("%s:l(%s,%d)", int, var, k),
                     sprintf("%s:l(%s, %d)", int, var, k))
      inter     <- .find_one(nm, int_cands)
    }
    low  <- c(low,  base)
    high <- c(high, inter)
  }
  list(low = low, high = high)
}

# Compute sum(beta_terms) with SE using chosen vcov
lincombo_sum <- function(mod, terms, vcov_arg = NULL){
  nm <- names(coef(mod))
  if (anyNA(terms)) {
    stop("Missing terms at positions: ", paste(which(is.na(terms)), collapse = ", "),
         "\nCheck names(coef(model)) to see how your lags/interactions are labeled.")
  }
  b  <- coef(mod)[terms]
  V  <- if (is.null(vcov_arg)) vcov(mod) else vcov(mod, vcov = vcov_arg)
  V  <- V[terms, terms, drop = FALSE]
  w  <- rep(1, length(terms))
  est <- sum(b)
  se  <- as.numeric(sqrt(t(w) %*% V %*% w))  # = sqrt(sum(V)) since w=1s
  z   <- est / se
  p   <- 2 * pnorm(abs(z), lower.tail = FALSE)
  ci  <- est + c(-1, 1) * 1.96 * se
  list(est = est, se = se, z = z, p = p, ci = ci, terms = terms)
}

# Wrapper to compute Low / High / Diff cumulative (0–2y)
cum_effects_0_2 <- function(mod, var = "cpi10", lags = 0:2, int = "HighCarbon", vcov_arg = NULL){
  nm   <- names(coef(mod))
  tt   <- get_terms_dynhet(nm, var = var, lags = lags, int = int)
  low  <- lincombo_sum(mod, tt$low, vcov_arg = vcov_arg)
  high <- lincombo_sum(mod, c(tt$low, tt$high), vcov_arg = vcov_arg)
  diff <- lincombo_sum(mod, tt$high, vcov_arg = vcov_arg)
  # return a neat data.frame
  tibble::tibble(
    group = c("Low (0–2y)", "High (0–2y)", "High − Low (0–2y)"),
    est   = c(low$est, high$est, diff$est),
    se    = c(low$se,  high$se,  diff$se),
    z     = c(low$z,   high$z,   diff$z),
    p     = c(low$p,   high$p,   diff$p),
    ci_lo = c(low$ci[1], high$ci[1], diff$ci[1]),
    ci_hi = c(low$ci[2], high$ci[2], diff$ci[2])
  )
}

# --------------------------
# 2) Fit BOTH models (PC and TOT) with lags interacted
# --------------------------
# Make sure HighCarbon is 0/1 numeric and data sorted
df_rq3 <- df_rq3 %>%
  mutate(HighCarbon = as.integer(HighCarbon)) %>%
  arrange(country, year)

# Per-capita
m_rq3_dyn_pc <- feols(
  log_co2_pc ~ cpi10*HighCarbon + l(cpi10,1)*HighCarbon + l(cpi10,2)*HighCarbon +
    lgdppc_c + ind_c + ren_c | country + year,
  data     = df_rq3,
  panel.id = ~ country + year,
  vcov     = ~ country + year  # keep two-way cluster here; DK via cum_effects_0_2 if desired
)

# Totals (adds population)
m_rq3_dyn_tot <- feols(
  log_co2 ~ cpi10*HighCarbon + l(cpi10,1)*HighCarbon + l(cpi10,2)*HighCarbon +
    log_pop + lgdppc_c + ind_c + ren_c | country + year,
  data     = df_rq3,
  panel.id = ~ country + year,
  vcov     = ~ country + year
)

# --------------------------
# 3) Cumulative (0–2y) effects using lincombo_sum for BOTH models
# --------------------------
# Use two-way cluster SEs (as estimated) -> vcov_arg = NULL
res_pc_tw  <- cum_effects_0_2(m_rq3_dyn_pc, var = "cpi10", lags = 0:2, int = "HighCarbon", vcov_arg = NULL)
res_tot_tw <- cum_effects_0_2(m_rq3_dyn_tot, var = "cpi10", lags = 0:2, int = "HighCarbon", vcov_arg = NULL)

# If you prefer DK SEs instead, pass vcov_arg = vDK:
# res_pc_dk  <- cum_effects_0_2(m_rq3_dyn_pc,  vcov_arg = vDK)
# res_tot_dk <- cum_effects_0_2(m_rq3_dyn_tot, vcov_arg = vDK)

print(res_pc_tw)
print(res_tot_tw)

# --------------------------
# 4) (Optional) Show models with etable, keep SEs below
# --------------------------
etable(
  m_rq3_dyn_pc, m_rq3_dyn_tot,
  se.below = TRUE,
  headers  = c("RQ3 Dynamic (Per-capita)", "RQ3 Dynamic (Totals + log_pop)")
)

```
### Tables
```{r}

# -------- Helper: nicely format "est (se)" ----------
fmt_es <- function(est, se, digits = 4) {
  sprintf(paste0("%.", digits, "f (%.", digits, "f)"), est, se)
}

# -------- Build the extra rows for cumulative effects ----------
row_lab_pc  <- "RQ3 Dynamic (Per-capita)"
row_lab_tot <- "RQ3 Dynamic (Totals + log_pop)"

add_rows_tbl <- tibble::tibble(
  term                         = c(" ", "Cumulative (0–2y): Low", "Cumulative (0–2y): High", "Cumulative (0–2y): High − Low"),
  `RQ3 Dynamic (Per-capita)`   = c("", fmt_es(res_pc_tw$est[res_pc_tw$group=="Low (0–2y)"],
                                              res_pc_tw$se [res_pc_tw$group=="Low (0–2y)"]),
                                   fmt_es(res_pc_tw$est[res_pc_tw$group=="High (0–2y)"],
                                              res_pc_tw$se [res_pc_tw$group=="High (0–2y)"]),
                                   fmt_es(res_pc_tw$est[res_pc_tw$group=="High − Low (0–2y)"],
                                              res_pc_tw$se [res_pc_tw$group=="High − Low (0–2y)"])
  ),
  `RQ3 Dynamic (Totals + log_pop)` = c("", fmt_es(res_tot_tw$est[res_tot_tw$group=="Low (0–2y)"],
                                                  res_tot_tw$se [res_tot_tw$group=="Low (0–2y)"]),
                                       fmt_es(res_tot_tw$est[res_tot_tw$group=="High (0–2y)"],
                                                  res_tot_tw$se [res_tot_tw$group=="High (0–2y)"]),
                                       fmt_es(res_tot_tw$est[res_tot_tw$group=="High − Low (0–2y)"],
                                                  res_tot_tw$se [res_tot_tw$group=="High − Low (0–2y)"])
  )
)

# --------- Optional: clean coefficient labels (auto-maps whatever names you have) ----------
nm_pc  <- names(coef(m_rq3_dyn_pc))
nm_tot <- names(coef(m_rq3_dyn_tot))
nm_all <- unique(c(nm_pc, nm_tot))

# Create a mapping that prints nice labels if it recognizes common patterns
pretty_map <- setNames(nm_all, nm_all)
pretty_map[grepl("^cpi10$", nm_all)]                    <- "CPI (t) / +10 €/t"
pretty_map[grepl("^l\\(cpi10,\\s*1\\)$", nm_all)]       <- "CPI (t−1)"
pretty_map[grepl("^l\\(cpi10,\\s*2\\)$", nm_all)]       <- "CPI (t−2)"
pretty_map[grepl("^cpi10:HighCarbon", nm_all)]          <- "× HighCarbon"
pretty_map[grepl("^HighCarbon:cpi10", nm_all)]          <- "× HighCarbon"
pretty_map[grepl("^l\\(cpi10,\\s*1\\):HighCarbon", nm_all)] <- "CPI (t−1) × HighCarbon"
pretty_map[grepl("^HighCarbon:l\\(cpi10,\\s*1\\)", nm_all)] <- "CPI (t−1) × HighCarbon"
pretty_map[grepl("^l\\(cpi10,\\s*2\\):HighCarbon", nm_all)] <- "CPI (t−2) × HighCarbon"
pretty_map[grepl("^HighCarbon:l\\(cpi10,\\s*2\\)", nm_all)] <- "CPI (t−2) × HighCarbon"

pretty_map[grepl("^lgdppc_c$", nm_all)] <- "log GDP pc (centered)"
pretty_map[grepl("^ind_c$", nm_all)]    <- "Industry share (centered)"
pretty_map[grepl("^ren_c$", nm_all)]    <- "Renewables (centered)"
pretty_map[grepl("^log_pop$", nm_all)]  <- "log Population"

# --------- Produce LaTeX table ----------
msummary(
  list(`RQ3 Dynamic (Per-capita)` = m_rq3_dyn_pc,
       `RQ3 Dynamic (Totals + log_pop)` = m_rq3_dyn_tot),
  output   = "latex",
  stars    = c('*' = .05, '**' = .01, '***' = .001),
  coef_map = pretty_map,
  gof_omit = 'IC|Log.Lik|Adj|AIC|BIC|RMSE',      # keep table tight
  add_rows = add_rows_tbl,
  notes = c(
    "+1 in CPI regressor = +10 €/t.",
    "Country & year FE in all models.",
    "SEs: two-way clustered by country & year; cumulative rows show est (SE)."
  ),
  title = "RQ3: Dynamic heterogeneity (0–2 years) across High/Low carbon-intensity countries"
)

```

```{r}

cum_tab <- dplyr::tibble(
  `Cumulative (0–2y)` = c("Low", "High", "High − Low"),
  `Per-capita`        = c(fmt_es(res_pc_tw$est[1],  res_pc_tw$se[1]),
                          fmt_es(res_pc_tw$est[2],  res_pc_tw$se[2]),
                          fmt_es(res_pc_tw$est[3],  res_pc_tw$se[3])),
  `Totals + log_pop`  = c(fmt_es(res_tot_tw$est[1], res_tot_tw$se[1]),
                          fmt_es(res_tot_tw$est[2], res_tot_tw$se[2]),
                          fmt_es(res_tot_tw$est[3], res_tot_tw$se[3]))
)

kbl(cum_tab, format = "html", booktabs = TRUE, align = "lcc",
    caption = "RQ3: Cumulative CPI effects (0–2y) by carbon-intensity group") |>
  kable_classic(full_width = FALSE, html_font = NULL) |>
  add_header_above(c(" " = 1, "Models" = 2)) |>
  footnote(
    general = "+1 in CPI regressor = +10 €/t. Country & year FE. SEs in parentheses.",
    threeparttable = TRUE
  )

```

#### tables unique

```{r}

# Pretty labels
pretty_map_rq3 <- c(
  "cpi10"                = "CPI (t) / +10 €/t",
  "HighCarbon"           = "HighCarbon (baseline absorbed)",  # absorbed by country FE
  "cpi10:HighCarbon"     = "CPI (t) × HighCarbon",
  "lgdppc_c"             = "log GDP pc (centered)",
  "ind_c"                = "Industry share (centered)",
  "ren_c"                = "Renewables (centered)",
  "log_pop"              = "log Population"
)

gof_omit <- "IC|Log.Lik|Adj|AIC|BIC|RMSE"
dir.create("tables", showWarnings = FALSE)

msummary(
  list(
    "RQ3 Direct: Per-capita (main)" = m_rq3,
    "RQ3 Direct: Totals + log_pop"  = m_rq3_direct_tot
  ),
  output    = "tables/rq3_direct_main.tex",
  stars     = c('*'=.05, '**'=.01, '***'=.001),
  statistic = "({std.error})",
  coef_map  = pretty_map_rq3,
  gof_omit  = gof_omit,
  notes = c(
    "Country & year fixed effects in all models.",
    "Standard errors two-way clustered by country and year.",
    "+1 in CPI regressor = +10 €/t."
  ),
  title = "RQ3 — Carbon-Intensity Heterogeneity: Direct Effect (Per-capita main; Totals robustness)"
)

```
```{r}
# Extend labels for lag and interaction terms (map both directions defensively)
pretty_map_rq3_dyn <- c(pretty_map_rq3,
  "l(cpi10,1)"                   = "CPI (t−1)",
  "l(cpi10,2)"                   = "CPI (t−2)",
  "l(cpi10,1):HighCarbon"        = "CPI (t−1) × HighCarbon",
  "HighCarbon:l(cpi10,1)"        = "CPI (t−1) × HighCarbon",
  "l(cpi10,2):HighCarbon"        = "CPI (t−2) × HighCarbon",
  "HighCarbon:l(cpi10,2)"        = "CPI (t−2) × HighCarbon"
)

msummary(
  list(
    "RQ3 Dynamic: Per-capita (main)" = m_rq3_dyn_pc,
    "RQ3 Dynamic: Totals + log_pop"  = m_rq3_dyn_tot
  ),
  output    = "tables/rq3_dynamic_main.tex",
  stars     = c('*'=.05, '**'=.01, '***'=.001),
  statistic = "({std.error})",
  coef_map  = pretty_map_rq3_dyn,
  gof_omit  = gof_omit,
  notes = c(
    "Country & year fixed effects in all models.",
    "Standard errors two-way clustered by country and year.",
    "Dynamic specs include CPI(t), CPI(t−1), and CPI(t−2) interacted with HighCarbon.",
    "+1 in CPI regressor = +10 €/t."
  ),
  title = "RQ3 — Carbon-Intensity Heterogeneity: Dynamic Effects (Per-capita main; Totals robustness)"
)

```

## RQ 4

```{r}
library(dplyr)
library(fixest)
library(modelsummary)
library(kableExtra)
library(tibble)

# =========================================================
# 0) Construct HighIncome (pre-period 2004–2006)
#    Use an UN-centered GDPpc log variable (e.g., "lgdppc").
# =========================================================
income_var <- "lgdppc_c"   # <-- change if yours is named differently

baseline_income <- analysis_sample %>%
  filter(year %in% 2004:2006) %>%
  group_by(country) %>%
  summarise(lgdppc_base = mean(.data[[income_var]], na.rm = TRUE), .groups = "drop") %>%
  mutate(HighIncome = ifelse(lgdppc_base > median(lgdppc_base, na.rm = TRUE), 1L, 0L))

df_rq4 <- analysis_sample %>%
  left_join(baseline_income, by = "country") %>%
  mutate(HighIncome = as.integer(HighIncome)) %>%
  arrange(country, year)

# =========================================================
# 1) DIRECT EFFECT (no lags), per-capita
# =========================================================
m_rq4_direct_pc <- feols(
  log_co2_pc ~ cpi10 * HighIncome + lgdppc_c + ind_c + ren_c | country + year,
  data = df_rq4,
  vcov = ~ country + year
)

# Optional: direct effect robustness on totals (adds population)
m_rq4_direct_tot <- feols(
  log_co2 ~ cpi10 * HighIncome + log_pop + lgdppc_c + ind_c + ren_c | country + year,
  data = df_rq4,
  vcov = ~ country + year
)

etable(m_rq4_direct_pc, m_rq4_direct_tot, se.below = TRUE,
       headers = c("RQ4 Direct: Per-capita", "RQ4 Direct: Totals + log_pop"))

# =========================================================
# 2) DYNAMIC EFFECT (lags up to 2), per-capita
# =========================================================
m_rq4_dyn_pc <- feols(
  log_co2_pc ~ cpi10*HighIncome + l(cpi10,1)*HighIncome + l(cpi10,2)*HighIncome +
    lgdppc_c + ind_c + ren_c | country + year,
  data     = df_rq4,
  panel.id = ~ country + year,
  vcov     = ~ country + year
)

# 3) DYNAMIC EFFECT robustness on totals
m_rq4_dyn_tot <- feols(
  log_co2 ~ cpi10*HighIncome + l(cpi10,1)*HighIncome + l(cpi10,2)*HighIncome +
    log_pop + lgdppc_c + ind_c + ren_c | country + year,
  data     = df_rq4,
  panel.id = ~ country + year,
  vcov     = ~ country + year
)

etable(m_rq4_dyn_pc, m_rq4_dyn_tot, se.below = TRUE,
       headers = c("RQ4 Dynamic: Per-capita", "RQ4 Dynamic: Totals + log_pop"))

# =========================================================
# 4) CUMULATIVE (0–2y) EFFECTS — helper functions (no wald())
# =========================================================
.find_one <- function(nm, candidates){
  hit <- candidates[candidates %in% nm]
  if (!length(hit)) NA_character_ else hit[1]
}

get_terms_dynhet <- function(nm, var = "cpi10", lags = 0:2, int = "HighIncome"){
  low <- high <- character(0)
  for (k in lags) {
    if (k == 0) {
      base  <- .find_one(nm, var)
      inter <- .find_one(nm, c(sprintf("%s:%s", var, int), sprintf("%s:%s", int, var)))
    } else {
      lag_cands <- c(sprintf("l(%s,%d)", var, k), sprintf("l(%s, %d)", var, k))
      base      <- .find_one(nm, lag_cands)
      int_cands <- c(sprintf("l(%s,%d):%s", var, k, int), sprintf("l(%s, %d):%s", var, k, int),
                     sprintf("%s:l(%s,%d)", int, var, k), sprintf("%s:l(%s, %d)", int, var, k))
      inter     <- .find_one(nm, int_cands)
    }
    low  <- c(low,  base)
    high <- c(high, inter)
  }
  list(low = low, high = high)
}

lincombo_sum <- function(mod, terms, vcov_arg = NULL){
  nm <- names(coef(mod))
  if (anyNA(terms)) stop("Missing terms: ", paste(which(is.na(terms)), collapse = ", "),
                         "\nCheck names(coef(model)).")
  b <- coef(mod)[terms]
  V <- if (is.null(vcov_arg)) vcov(mod) else vcov(mod, vcov = vcov_arg)
  V <- V[terms, terms, drop = FALSE]
  w <- rep(1, length(terms))
  est <- sum(b)
  se  <- as.numeric(sqrt(t(w) %*% V %*% w))
  z   <- est / se
  p   <- 2 * pnorm(abs(z), lower.tail = FALSE)
  ci  <- est + c(-1, 1) * 1.96 * se
  list(est = est, se = se, z = z, p = p, ci = ci, terms = terms)
}

cum_effects_0_2 <- function(mod, var = "cpi10", lags = 0:2, int = "HighIncome", vcov_arg = NULL){
  nm <- names(coef(mod))
  tt <- get_terms_dynhet(nm, var = var, lags = lags, int = int)
  low  <- lincombo_sum(mod, tt$low, vcov_arg = vcov_arg)
  high <- lincombo_sum(mod, c(tt$low, tt$high), vcov_arg = vcov_arg)
  diff <- lincombo_sum(mod, tt$high, vcov_arg = vcov_arg)
  tibble(
    group = c("Low (0–2y)", "High (0–2y)", "High − Low (0–2y)"),
    est   = c(low$est, high$est, diff$est),
    se    = c(low$se,  high$se,  diff$se),
    z     = c(low$z,   high$z,   diff$z),
    p     = c(low$p,   high$p,   diff$p),
    ci_lo = c(low$ci[1], high$ci[1], diff$ci[1]),
    ci_hi = c(low$ci[2], high$ci[2], diff$ci[2])
  )
}

# Compute cumulative effects (two-way clustered SEs from the models)
res_rq4_pc  <- cum_effects_0_2(m_rq4_dyn_pc,  var = "cpi10", int = "HighIncome")
res_rq4_tot <- cum_effects_0_2(m_rq4_dyn_tot, var = "cpi10", int = "HighIncome")

res_rq4_pc
res_rq4_tot

```
### Tables
```{r app RQ4}
fmt_es <- function(est, se, digits = 4) sprintf(paste0("%.",digits,"f (%.",digits,"f)"), est, se)

cum_tab_rq4 <- tibble(
  `Cumulative (0–2y)` = c("Low", "High", "High − Low"),
  `Per-capita`        = c(fmt_es(res_rq4_pc$est[1],  res_rq4_pc$se[1]),
                          fmt_es(res_rq4_pc$est[2],  res_rq4_pc$se[2]),
                          fmt_es(res_rq4_pc$est[3],  res_rq4_pc$se[3])),
  `Totals + log_pop`  = c(fmt_es(res_rq4_tot$est[1], res_rq4_tot$se[1]),
                          fmt_es(res_rq4_tot$est[2], res_rq4_tot$se[2]),
                          fmt_es(res_rq4_tot$est[3], res_rq4_tot$se[3]))
)

kbl(cum_tab_rq4, format = "html", booktabs = TRUE, align = "lcc",
    caption = "RQ4: Cumulative CPI effects (0–2y) by income group") |>
  kable_classic(full_width = FALSE) |>
  add_header_above(c(" " = 1, "Models" = 2)) |>
  footnote(
    general = "+1 in CPI regressor = +10 €/t. Country & year FE. SEs: two-way clustered by country & year. Values are estimates with SE in parentheses.",
    threeparttable = TRUE
  )

```
Direct effect: per-capita (main) + totals (robustness)

```{r}
library(modelsummary)
library(tibble)
library(stringr)

# Pretty labels for key regressors
pretty_map <- c(
  "cpi10"                              = "CPI (t) / +10 €/t",
  "HighIncome"                         = "HighIncome (baseline absorbed)", # FE absorbs its main effect
  "cpi10:HighIncome"                   = "CPI (t) × HighIncome",
  "lgdppc_c"                           = "log GDP pc (centered)",
  "ind_c"                              = "Industry share (centered)",
  "ren_c"                              = "Renewables (centered)",
  "log_pop"                            = "log Population"
)

# Show clean GOF; hide IC etc.
gof_omit <- "IC|Log.Lik|Adj|AIC|BIC|RMSE"

# Build LaTeX table
dir.create("tables", showWarnings = FALSE)

msummary(
  list(
    "RQ4 Direct: Per-capita (main)"    = m_rq4_direct_pc,
    "RQ4 Direct: Totals + log_pop"     = m_rq4_direct_tot
  ),
  output   = "tables/rq4_direct_main.tex",
  stars    = c('*' = .05, '**' = .01, '***' = .001),
  statistic = "({std.error})",
  coef_map = pretty_map,
  gof_omit = gof_omit,
  notes = c(
    "Country & year fixed effects in all models.",
    "Standard errors two-way clustered by country and year.",
    "+1 in CPI regressor = +10 €/t."
  ),
  title = "RQ4 — Income Heterogeneity: Direct Effect (Per-capita main; Totals robustness)"
)

```

Dynamic effect (lags up to 2): per-capita (main) + totals (robustness)  
```{r}
# Extend pretty labels for lagged terms
pretty_map_dyn <- c(pretty_map,
  "l(cpi10,1)"                         = "CPI (t−1)",
  "l(cpi10,2)"                         = "CPI (t−2)",
  # Interactions (order can vary; map both directions defensively)
  "l(cpi10,1):HighIncome"              = "CPI (t−1) × HighIncome",
  "HighIncome:l(cpi10,1)"              = "CPI (t−1) × HighIncome",
  "l(cpi10,2):HighIncome"              = "CPI (t−2) × HighIncome",
  "HighIncome:l(cpi10,2)"              = "CPI (t−2) × HighIncome"
)

msummary(
  list(
    "RQ4 Dynamic: Per-capita (main)"   = m_rq4_dyn_pc,
    "RQ4 Dynamic: Totals + log_pop"    = m_rq4_dyn_tot
  ),
  output   = "tables/rq4_dynamic_main.tex",
  stars    = c('*' = .05, '**' = .01, '***' = .001),
  statistic = "({std.error})",
  coef_map = pretty_map_dyn,
  gof_omit = gof_omit,
  notes = c(
    "Country & year fixed effects in all models.",
    "Standard errors two-way clustered by country and year.",
    "Dynamic specs include CPI(t), CPI(t−1), CPI(t−2) interacted with HighIncome.",
    "+1 in CPI regressor = +10 €/t."
  ),
  title = "RQ4 — Income Heterogeneity: Dynamic Effects (Per-capita main; Totals robustness)"
)

```

# Graphs 

## RQ3 4

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(fixest)
library(stringr)

# ---- unified plot theme ----
theme_paper <- function() {
  theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      plot.title = element_text(face = "bold"),
      legend.position = "top"
    )
}

# ---- name finder (handles spaces in l( , ) and : order) ----
.find_one <- function(nm, candidates){
  hit <- candidates[candidates %in% nm]
  if (!length(hit)) NA_character_ else hit[1]
}

# ---- collect base & interaction terms for horizons 0:K ----
collect_terms <- function(nm, var = "cpi10", K = 2, int = "HighCarbon"){
  tibble(h = 0:K) %>%
    rowwise() %>%
    mutate(
      term_base = if (h == 0) .find_one(nm, var) else
                  .find_one(nm, c(sprintf("l(%s,%d)", var, h), sprintf("l(%s, %d)", var, h))),
      term_int  = if (h == 0) .find_one(nm, c(sprintf("%s:%s", var, int), sprintf("%s:%s", int, var))) else
                  .find_one(nm, c(sprintf("l(%s,%d):%s", var, h, int),
                                  sprintf("l(%s, %d):%s", var, h, int),
                                  sprintf("%s:l(%s,%d)", int, var, h),
                                  sprintf("%s:l(%s, %d)", int, var, h)))
    ) %>%
    ungroup()
}

# ---- extract horizon-by-horizon effects (Low = base; High = base+int) ----
dyn_effects_df <- function(mod, var = "cpi10", K = 2, int = "HighCarbon", vcov_arg = ~ country + year) {
  nm   <- names(coef(mod))
  V    <- vcov(mod, vcov = vcov_arg)
  terms <- collect_terms(nm, var, K, int)

  # build rows for Low and High at each horizon
  df_list <- lapply(1:nrow(terms), function(i){
    tb <- terms[i,]
    # LOW: just base term
    low_terms  <- na.omit(c(tb$term_base))
    # HIGH: base + interaction
    high_terms <- na.omit(c(tb$term_base, tb$term_int))

    # helper to get est/se for a set of terms with weight 1
    est_se <- function(trms){
      b <- coef(mod)[trms]
      VV <- V[trms, trms, drop = FALSE]
      est <- sum(b)
      se  <- sqrt(sum(VV)) # weights 1
      c(est = est, se = se)
    }
    c_low  <- est_se(low_terms)
    c_high <- est_se(high_terms)

    tibble(
      h = tb$h,
      group = c("Low", "High"),
      est = c(c_low["est"], c_high["est"]),
      se  = c(c_low["se"],  c_high["se"])
    )
  })
  bind_rows(df_list) %>%
    mutate(ci_lo = est - 1.96*se, ci_hi = est + 1.96*se)
}

# ---- cumulative (0:K) Low/High/Diff ----
cum_0K_df <- function(mod, var = "cpi10", K = 2, int = "HighCarbon", vcov_arg = ~ country + year){
  nm <- names(coef(mod))
  V  <- vcov(mod, vcov = vcov_arg)
  terms <- collect_terms(nm, var, K, int)

  # assemble vectors of term names
  low_terms  <- na.omit(terms$term_base)
  high_add   <- na.omit(terms$term_int)
  high_terms <- na.omit(c(low_terms, high_add))
  diff_terms <- na.omit(high_add)

  est_se <- function(trms){
    b <- coef(mod)[trms]
    VV <- V[trms, trms, drop = FALSE]
    est <- sum(b); se <- sqrt(sum(VV))
    c(est = est, se = se)
  }
  LOW  <- est_se(low_terms)
  HIGH <- est_se(high_terms)
  DIFF <- est_se(diff_terms)

  tibble(
    group = c("Low (0–2y)", "High (0–2y)", "High − Low (0–2y)"),
    est   = c(LOW["est"], HIGH["est"], DIFF["est"]),
    se    = c(LOW["se"],  HIGH["se"],  DIFF["se"])
  ) %>% mutate(ci_lo = est - 1.96*se, ci_hi = est + 1.96*se)
}

```

rq 3- dynamic effect horizons

```{r}
df_dyn_rq3 <- dyn_effects_df(m_rq3_dyn_pc, var = "cpi10", K = 2, int = "HighCarbon", vcov_arg = ~ country + year)

gg_rq3_dyn <- ggplot(df_dyn_rq3, aes(x = h, y = est, color = group)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi, fill = group), alpha = 0.15, color = NA) +
  geom_point(size = 2) +
  geom_line() +
  scale_x_continuous(breaks = 0:2, labels = c("t","t−1","t−2")) +
  labs(
    title = "RQ3: Dynamic CPI effect by horizon (per-capita)",
    x = "Horizon",
    y = "Effect on log CO₂ per capita",
    color = "", fill = ""
  ) +
  theme_paper()

gg_rq3_dyn

```
rq 3 - cumulative effects

```{r}
df_cum_rq3 <- cum_0K_df(m_rq3_dyn_pc, var = "cpi10", K = 2, int = "HighCarbon", vcov_arg = ~ country + year)

gg_rq3_cum <- ggplot(df_cum_rq3, aes(x = group, y = est)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_col(width = 0.6) +
  geom_errorbar(aes(ymin = ci_lo, ymax = ci_hi), width = 0.15) +
  labs(
    title = "RQ3: Cumulative CPI effect (0–2 years)",
    x = "", y = "Cumulative effect on log CO2 per capita"
  ) +
  theme_paper()

gg_rq3_cum

```
rq 3- direct effect forecast

```{r}
# pull CPI main & interaction with HighCarbon from the direct model m_rq3
summ <- summary(m_rq3, vcov = ~ country + year)
tab  <- broom::tidy(summ)

forest_rq3 <- tab %>%
  filter(term %in% c("cpi10", "cpi10:HighCarbon", "HighCarbon:cpi10")) %>%
  mutate(term = recode(term,
                       "cpi10" = "CPI (t)",
                       "cpi10:HighCarbon" = "CPI (t) × HighCarbon",
                       "HighCarbon:cpi10" = "CPI (t) × HighCarbon")) %>%
  mutate(term = factor(term, levels = rev(c("CPI (t)", "CPI (t) × HighCarbon"))))

gg_rq3_direct <- ggplot(forest_rq3, aes(y = term, x = estimate)) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = estimate - 1.96*std.error,
                     xmax = estimate + 1.96*std.error), height = 0.2) +
  labs(
    title = "RQ3: Direct effects (per-capita)",
    x = "Coefficient on log CO2 per capita", y = ""
  ) +
  theme_paper()

gg_rq3_direct

```
rq 4 - dynamic effect horizons

```{r not working}
# Dynamic by horizon
df_dyn_rq4 <- dyn_effects_df(m_rq4_dyn_pc, var = "cpi10", K = 2, int = "HighIncome", vcov_arg = ~ country + year)

gg_rq4_dyn <- ggplot(df_dyn_rq4, aes(x = h, y = est, color = group)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi, fill = group), alpha = 0.15, color = NA) +
  geom_point(size = 2) +
  geom_line() +
  scale_x_continuous(breaks = 0:2, labels = c("t","t−1","t−2")) +
  labs(
    title = "RQ4: Dynamic CPI effect by horizon (per-capita)",
    x = "Horizon",
    y = "Effect on log CO₂ per capita",
    color = "", fill = ""
  ) +
  theme_paper()

# Cumulative bars
df_cum_rq4 <- cum_0K_df(m_rq4_dyn_pc, var = "cpi10", K = 2, int = "HighIncome", vcov_arg = ~ country + year)

gg_rq4_cum <- ggplot(df_cum_rq4, aes(x = group, y = est)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_col(width = 0.6) +
  geom_errorbar(aes(ymin = ci_lo, ymax = ci_hi), width = 0.15) +
  labs(
    title = "RQ4: Cumulative CPI effect (0–2 years)",
    x = "", y = "Cumulative effect on log CO₂ per capita"
  ) +
  theme_paper()

# Direct-effect forest
summ4 <- summary(m_rq4_direct_pc, vcov = ~ country + year)
tab4  <- broom::tidy(summ4)

forest_rq4 <- tab4 %>%
  filter(term %in% c("cpi10", "cpi10:HighIncome", "HighIncome:cpi10")) %>%
  mutate(term = recode(term,
                       "cpi10" = "CPI (t)",
                       "cpi10:HighIncome" = "CPI (t) × HighIncome",
                       "HighIncome:cpi10" = "CPI (t) × HighIncome")) %>%
  mutate(term = factor(term, levels = rev(c("CPI (t)", "CPI (t) × HighIncome"))))

gg_rq4_direct <- ggplot(forest_rq4, aes(y = term, x = estimate)) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = estimate - 1.96*std.error,
                     xmax = estimate + 1.96*std.error), height = 0.2) +
  labs(
    title = "RQ4: Direct effects (per-capita)",
    x = "Coefficient on log CO₂ per capita", y = ""
  ) +
  theme_paper()

```
```{r}
library(dplyr)
library(ggplot2)
library(stringr)

cpi_plot <- analysis_sample %>%
  mutate(year = as.integer(year),
         cpi_eur = 10 * cpi10) %>%
  filter(year >= 2005, year <= 2023) %>%
  arrange(country, year)

```

```{r}
p_facets <- ggplot(cpi_plot, aes(x = year, y = cpi_eur)) +
  geom_hline(yintercept = 0, linetype = 2, linewidth = 0.2) +
  geom_line(linewidth = 0.5) +
  facet_wrap(~ country, ncol = 6) +  # adjust columns to fit your layout
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020)) +
  labs(title = "Carbon Price Index by Country (€/t)",
       x = NULL, y = "CPI (€/t)") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))
p_facets

```
If some countries have tiny variance and look flat, switch to scales = "free_y" to emphasize shapes (but lose cross-panel comparability):
```{r}
p_facets_free <- p_facets + facet_wrap(~ country, ncol = 6, scales = "free_y")
p_facets_free
```

```{r}
highlight <- c("Germany","France","Netherlands","Sweden")  # or ISO codes matching your `country`

cpi_highlight <- cpi_plot %>%
  mutate(group = if_else(country %in% highlight, "Highlighted","Other"))

p_high <- ggplot(cpi_highlight, aes(year, cpi_eur, group = country)) +
  geom_line(data = ~ filter(.x, group=="Other"),
            alpha = 0.35, linewidth = 0.4, color = "grey50") +
  geom_line(data = ~ filter(.x, group=="Highlighted"),
            linewidth = 0.9) +
  labs(title = "Carbon Price Index (€/t): highlight vs. others",
       x = NULL, y = "CPI (€/t)") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"))
p_high

```
```{r}
library(ggplot2)
library(dplyr)

# ---- Data prep ----
cpi_plot <- analysis_sample %>%
  mutate(year = as.integer(year),
         cpi_eur = 10 * cpi10) %>%
  filter(year >= 2005, year <= 2023) %>%
  arrange(country, year)

# ---- Define EU ETS phases with higher-contrast greens ----
phase_rects <- tibble::tibble(
  xmin  = c(2005, 2008, 2013, 2021),
  xmax  = c(2007.99, 2012.99, 2020.99, 2023.99),
  phase = c("Phase I", "Phase II", "Phase III", "Phase IV"),
  fill  = c("#E5F5E0", "#A1D99B", "#41AB5D", "#006D2C")
)

# ---- Plot ----
p_cpi_ets_contrast <- ggplot() +
  geom_rect(data = phase_rects,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = phase),
            alpha = 0.25, color = NA) +     # higher alpha for stronger contrast
  geom_line(data = cpi_plot, aes(year, cpi_eur, group = country),
            linewidth = 0.5, color = "black") +
  facet_wrap(~ country, ncol = 6) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020)) +
  scale_fill_manual(values = phase_rects$fill, name = "EU ETS Phase") +
  labs(title = "Carbon Price Index by Country with EU ETS Phases (2005–2023)",
       x = NULL, y = "CPI (€/t)") +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold"),
    plot.title = element_text(face = "bold"),
    legend.position = "bottom",
    legend.box = "horizontal"
  )

p_cpi_ets_contrast


```

```{r}
dir.create("figs", showWarnings = FALSE)
ggsave("figs/cpi_by_country_phases_greencontrast.pdf", p_cpi_ets_contrast)
ggsave("figs/cpi_by_country_phases_greencontrast.png", p_cpi_ets_contrast, width = 8, height = 6, dpi = 300)

```

```{r}
library(dplyr)
library(ggplot2)

# Data prep
cpi_lines <- analysis_sample %>%
  mutate(year = as.integer(year),
         cpi_eur = 10 * cpi10) %>%
  filter(year >= 2005, year <= 2023) %>%
  arrange(country, year)

# 1) Simple multi-country colored lines (legend shown)
p_cpi_lines <- ggplot(cpi_lines, aes(x = year, y = cpi_eur, color = country, group = country)) +
  geom_line(linewidth = 0.7, alpha = 0.9) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020, 2023)) +
  labs(title = "Carbon Price Index by Country (2005–2023)",
       x = NULL, y = "CPI (€/t)", color = "Country") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"),
        legend.position = "right")
p_cpi_lines

# If you prefer a colorblind-friendly palette:
# + scale_color_viridis_d(option = "C", end = 0.95)

# 2) End-of-line labels instead of a big legend (nice for many countries)
library(ggrepel)
last_pts <- cpi_lines %>%
  group_by(country) %>%
  filter(year == max(year)) %>%
  ungroup()

p_cpi_labeled <- ggplot(cpi_lines, aes(year, cpi_eur, group = country)) +
  geom_line(aes(color = country), linewidth = 0.7, alpha = 0.9, show.legend = FALSE) +
  geom_point(data = last_pts, aes(color = country), size = 1.5, show.legend = FALSE) +
  ggrepel::geom_text_repel(
    data = last_pts,
    aes(label = country, color = country),
    nudge_x = 0.4, direction = "y", hjust = 0, segment.alpha = 0.4, show.legend = FALSE, size = 3
  ) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020, 2023), limits = c(2005, 2024.5)) +
  labs(title = "Carbon Price Index by Country (2005–2023)",
       x = NULL, y = "CPI (€/t)") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold"),
        plot.margin = margin(5.5, 40, 5.5, 5.5)) # space for labels on right
p_cpi_labeled

# 3) Faceted version if lines are too crowded (common y-axis)
p_cpi_facets <- ggplot(cpi_lines, aes(year, cpi_eur, group = country)) +
  geom_line(linewidth = 0.6) +
  facet_wrap(~ country, ncol = 6) +  # adjust columns to fit layout
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020, 2023)) +
  labs(title = "Carbon Price Index by Country (Facets, 2005–2023)",
       x = NULL, y = "CPI (€/t)") +
  theme_minimal(base_size = 10) +
  theme(panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))
p_cpi_facets

# Save PDFs for LaTeX
dir.create("figs", showWarnings = FALSE)
ggsave("figs/cpi_lines_all.pdf", p_cpi_lines,   width = 7.2, height = 4.6, device = cairo_pdf)
ggsave("figs/cpi_lines_labels.pdf", p_cpi_labeled, width = 7.6, height = 5.0, device = cairo_pdf)
ggsave("figs/cpi_lines_facets.pdf", p_cpi_facets, width = 7.5, height = 9.0, device = cairo_pdf)

```

```{r}

# --- Prep: CPI in €/t, 2005–2023, and sort countries by last CPI (for nicer label stacking)
cpi_lines <- analysis_sample %>%
  mutate(year = as.integer(year),
         cpi_eur = 10 * cpi10) %>%
  filter(year >= 2005, year <= 2023) %>%
  arrange(country, year)

last_pts <- cpi_lines %>%
  group_by(country) %>%
  filter(year == max(year)) %>%
  ungroup() %>%
  arrange(desc(cpi_eur))

# Fix factor order by last CPI so colors/labels are deterministic & stacked
cpi_lines <- cpi_lines %>%
  mutate(country = factor(country, levels = unique(last_pts$country)))
last_pts   <- last_pts   %>%
  mutate(country = factor(country, levels = levels(cpi_lines$country)))

# --- Colorblind-safe qualitative palette (scales with N automatically)
# Alternatives: "Set 2", "Set 3", "Dark 2"; "Dark 3" gives good contrast on white.
cols <- setNames(hcl.colors(n = nlevels(cpi_lines$country), palette = "Dark 3"),
                 levels(cpi_lines$country))

# --- (Optional) EU ETS phase shading — set use_phases <- TRUE to enable
use_phases <- FALSE
phase_rects <- tibble::tibble(
  xmin  = c(2005, 2008, 2013, 2021),
  xmax  = c(2007.99, 2012.99, 2020.99, 2023.99),
  phase = c("Phase I","Phase II","Phase III","Phase IV")
)

base_plot <- ggplot() +
  { if (use_phases) geom_rect(data = phase_rects,
                              aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf),
                              inherit.aes = FALSE, alpha = 0.10) } +
  geom_line(data = cpi_lines,
            aes(x = year, y = cpi_eur, group = country, color = country),
            linewidth = 0.8, alpha = 0.95, show.legend = FALSE) +
  geom_point(data = last_pts,
             aes(x = year, y = cpi_eur, color = country),
             size = 1.6, show.legend = FALSE) +
  ggrepel::geom_text_repel(
    data = last_pts,
    aes(x = year, y = cpi_eur, label = country, color = country),
    nudge_x = 0.6, direction = "y", hjust = 0,
    min.segment.length = 0, segment.alpha = 0.4, segment.size = 0.3,
    box.padding = 0.15, point.padding = 0.15, max.overlaps = Inf,
    show.legend = FALSE, size = 3.0
  ) +
  scale_color_manual(values = cols) +
  scale_x_continuous(breaks = c(2005, 2006, 2007, 2008, 2009, 2010,2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023),
                     limits = c(2005, 2024.8)) +  # room on the right for labels
  labs(
    title = "Carbon Price Index by Country (2005–2023)",
    x = NULL, y = "CPI (€/t)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    plot.title = element_text(face = "bold"),
    plot.margin = margin(5.5, 60, 5.5, 5.5),  # extra right margin for labels
    legend.position = "none",
    axis.line.x = element_line(linewidth = 0.1, colour = "grey40"),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
  ) +
  coord_cartesian(clip = "off")

p_cpi_labeled_better <- base_plot
p_cpi_labeled_better



```

```{r}
# --- Phase boundaries & labels ---
phase_boundaries <- c(2005, 2008, 2013, 2021)
phase_labels <- data.frame(
  x = c(2006.5, 2010, 2016.5, 2022),    # midpoints for labels
  lab = c("Phase I", "Phase II", "Phase III", "Phase IV")
)

p_cpi_labeled_etslines <- base_plot +
  # vertical lines for phase boundaries
  geom_vline(xintercept = phase_boundaries,
             linetype = c("solid","dashed","dashed","dashed"),
             linewidth = 0.3, color = "grey35") +
  # phase labels at top (adjust y if needed)
  annotate("text",
           x = phase_labels$x, y = Inf, vjust = 1.6,
           label = phase_labels$lab,
           size = 3.2, fontface = "bold", color = "grey20") +
  # tiny buffer so labels fit
  coord_cartesian(clip = "off", xlim = c(2005, 2024.9))

p_cpi_labeled_etslines

ggsave("figs/cpi_lines_labeled_etslines.pdf", p_cpi_labeled_etslines)
```


# Analysis
## Collinearity diagnostics by Imai and Kim

```{r data prep}
# Prepare CPI variable: remove common year shocks (FE) & standardize within-country
m_chk <- lm(analysis_sample$cpi10 ~ factor(country) + factor(year), data=analysis_sample)
1 - summary(m_chk)$r.squared   # share of variance left after FE


# Build “idiosyncratic CPI” (remove common year shocks)
analysis_sample$cpi10_id <- fixest::feols(cpi10 ~ 1 | year, data=analysis_sample)$residuals


# Standardize within-country for stability & interpretability
analysis_sample <- analysis_sample |>
  dplyr::group_by(country) |>
  dplyr::mutate(cpi10_w = as.numeric(scale(cpi10, center=TRUE, scale=TRUE))) |>
  dplyr::ungroup()

# Scale for interpretability (effect per 1 SD of idiosyncratic CPI)
analysis_sample$cpi10_id_sd <- as.numeric(scale(analysis_sample$cpi10_id))

# ensure time is numeric
analysis_sample <- analysis_sample |>
  dplyr::mutate(year = as.integer(year))

# centered year for linear trends
analysis_sample <- analysis_sample %>%
  mutate(cyear = year - min(year, na.rm = TRUE))

```

With ~8.6% within-year-and-country variance, year FE are soaking up most ETS-driven swings, but there’s still signal from country-specific deviations (coverage changes, taxes, timing).

Expect larger SEs and sensitivity to over-parameterization (too many lags/interactions).

## RQ1: Baseline and robustness specs with Driscoll-Kraay SEs
```{r rq1 DK}
# --- (1) Baseline  -------------------------------------------
m_rq1_base <- feols(
  log_co2_pc ~ cpi10_id_sd + lgdppc_c + ind_c + ren_c | country + year,
  data     = analysis_sample,
  vcov     = "DK",
  panel.id = ~ country + year
)

# --- (2) + Country-specific linear trends --------------------------------
m_rq1_trend <- feols(
  log_co2_pc ~ cpi10_id_sd + lgdppc_c + ind_c + ren_c + cyear:country | country + year,
  data     = analysis_sample,
  vcov     = "DK",
  panel.id = ~ country + year
)

# --- (3) Total emissions as DV -------------------------------------------
# If you don't have it yet, create log total emissions:
m_rq1_total <- feols(
  log_co2 ~ cpi10_id_sd + lgdppc_c + ind_c + ren_c | country + year,
  data     = analysis_sample,
  vcov     = "DK",
  panel.id = ~ country + year
)

# --- (4) No renewables control -------------------------------------------
m_rq1_noren <- feols(
  log_co2_pc ~ cpi10_id_sd + lgdppc_c + ind_c | country + year,
  data     = analysis_sample,
  vcov     = "DK",
  panel.id = ~ country + year
)

# =============
# Wald tests 
# =============

# H0: CPI effect = 0 (each spec)
wald_1df <- function(m, name, vc = vcov(m)){
  b <- coef(m)[name]; V <- vc[name, name]
  W <- (b^2)/V
  p <- 1 - pchisq(W, df = 1)
  c(stat = W, p.value = p)
}

# Baseline: m_rq1_base

wald_1df(m_rq1_base, "cpi10_id_sd", vc = vcov(m_rq1_base))          # DK

vc_cl2_base <- sandwich::vcovCL(m_rq1_base, cluster = analysis_sample$country)
wald_1df(m_rq1_base, "cpi10_id_sd", vc = vc_cl2_base) #cluster


# Trends: m_rq1_trend
wald_1df(m_rq1_trend, "cpi10_id_sd", vc = vcov(m_rq1_trend))          # DK

vc_cl2 <- sandwich::vcovCL(m_rq1_trend, cluster = analysis_sample$country)
wald_1df(m_rq1_trend, "cpi10_id_sd", vc = vc_cl2) #cluster

# Total emissions: m_rq1_total
wald_1df(m_rq1_total, "cpi10_id_sd", vc = vcov(m_rq1_total))          # DK
vc_cl2_total <- sandwich::vcovCL(m_rq1_total, cluster = analysis_sample$country)
wald_1df(m_rq1_total, "cpi10_id_sd", vc = vc_cl2_total) #cluster

# No renewables: m_rq1_noren
wald_1df(m_rq1_noren, "cpi10_id_sd", vc = vcov(m_rq1_noren))          # DK
vc_cl2_noren <- sandwich::vcovCL(m_rq1_noren, cluster = analysis_sample$country)
wald_1df(m_rq1_noren, "cpi10_id_sd", vc = vc_cl2_noren) #cluster


# ====================
# Etable

etable(m_rq1_base, m_rq1_trend, m_rq1_total, m_rq1_noren,
         dict = c(
    cpi10_id_sd = "CPI (idiosyncratic, 1 SD)",
    lgdppc_c    = "log GDP pc (centered)",
    ind_c       = "Industry share",
    ren_c       = "Renewables share",
    `cyear:country` = "Country trends"
  ))

# =========================
# Translate CPI effect to "% for +10 CPI points" (each spec)
# =========================
sd_id <- sd(analysis_sample$cpi10_id, na.rm = TRUE)   # SD of *idiosyncratic* CPI in original units
scale_to_pct <- function(model, k = 10){
  b  <- coef(model)["cpi10_id_sd"]
  ci <- confint(model)["cpi10_id_sd", ]
  c(point = 100 * b  * (k / sd_id),
    lo    = 100 * ci[1] * (k / sd_id),
    hi    = 100 * ci[2] * (k / sd_id))
}

pct_base  <- scale_to_pct(m_rq1_base)
pct_trend <- scale_to_pct(m_rq1_trend)
pct_total <- scale_to_pct(m_rq1_total)
pct_noren <- scale_to_pct(m_rq1_noren)

pct_base; pct_trend; pct_total; pct_noren

```


```{r rq1 wald table}

# --- Helper: 1-df Wald test given a VCOV matrix ----------------------------
wald_1df <- function(m, name, vc = vcov(m)){
  b <- coef(m)[name]
  V <- vc[name, name]
  W <- (b^2) / V
  p <- 1 - pchisq(W, df = 1)
  c(stat = as.numeric(W), p.value = as.numeric(p))
}

# --- Your four models (already estimated) ----------------------------------
# m_rq1_base, m_rq1_trend, m_rq1_total, m_rq1_noren

mods <- list(
  "Baseline"        = m_rq1_base,
  "+ Country trends"= m_rq1_trend,
  "Total emissions" = m_rq1_total,
  "No renewables"   = m_rq1_noren
)

target <- "cpi10_id_sd"

# --- Build results frame ----------------------------------------------------
res <- lapply(names(mods), function(lbl){
  m <- mods[[lbl]]

  # DK vcov: comes from the model (since you estimated with vcov="DK")
  vc_dk <- stats::vcov(m)

  # Clustered-by-country vcov: supply a vector from the model frame
  mf <- model.frame(m)
  vc_cl <- sandwich::vcovCL(m, cluster = mf[["country"]])

  # Coef & SEs
  b    <- unname(coef(m)[target])
  se_dk <- sqrt(vc_dk[target, target])
  se_cl <- sqrt(vc_cl[target, target])

  # Wald tests
  w_dk <- wald_1df(m, target, vc = vc_dk)
  w_cl <- wald_1df(m, target, vc = vc_cl)

  # Fit stats
  N   <- fixest::fitstat(m, "n")
  wr2 <- fixest::fitstat(m, "wr2")

  data.frame(
    Specification    = lbl,
    Beta_CPI         = b,
    SE_DK            = se_dk,
    Wald_DK          = w_dk["stat"],
    p_DK             = w_dk["p.value"],
    SE_Cluster       = se_cl,
    Wald_Cluster     = w_cl["stat"],
    p_Cluster        = w_cl["p.value"],
    N                = as.integer(N),
    Within_R2        = as.numeric(wr2),
    check.names = FALSE
  )
}) %>% bind_rows()

# Pretty rounding
res_print <- res %>%
  mutate(
    Beta_CPI     = sprintf("%.4f", Beta_CPI),
    SE_DK        = sprintf("(%.4f)", SE_DK),
    Wald_DK      = sprintf("%.3f", Wald_DK),
    p_DK         = sprintf("%.3f", p_DK),
    SE_Cluster   = sprintf("(%.4f)", SE_Cluster),
    Wald_Cluster = sprintf("%.3f", Wald_Cluster),
    p_Cluster    = sprintf("%.3f", p_Cluster),
    Within_R2    = sprintf("%.3f", Within_R2)
  )

sd_id <- sd(analysis_sample$cpi10_id, na.rm = TRUE)

scale_to_pct <- function(b, sd_id, k = 10){
  100 * as.numeric(b) * (k / sd_id)
}

res_print <- res %>%
  mutate(
    Beta_CPI     = sprintf("%.4f", Beta_CPI),
    SE_DK        = sprintf("(%.4f)", SE_DK),
    Wald_DK      = sprintf("%.3f", Wald_DK),
    p_DK         = sprintf("%.3f", p_DK),
    SE_Cluster   = sprintf("(%.4f)", SE_Cluster),
    Wald_Cluster = sprintf("%.3f", Wald_Cluster),
    p_Cluster    = sprintf("%.3f", p_Cluster),
    Within_R2    = sprintf("%.3f", Within_R2),
    PctEff_plus10CPI = sprintf("%.2f%%", scale_to_pct(res$Beta_CPI, sd_id))
  )

# Output: set format = "latex" or "html"
out_format <- "html"
cap <- "RQ1 — CPI (idiosyncratic, 1 SD): Coefficients, SEs, and Wald tests (H0: β=0) with DK and Country-Clustered VCOV."

kbl(res_print,
    format   = out_format,
    booktabs = TRUE,
    caption  = cap,
    align    = "lrrrrrrrrrr") %>%    # 1 'l' + 10 'r' for 11 columns
  kable_styling(full_width = FALSE, position = "center") %>%
  add_header_above(c(" " = 2,
                     "DK SE / Wald" = 3,
                     "Clustered (country) SE / Wald" = 3,
                     "Fit" = 2,
                     "% effect (+10 CPI)" = 1))



```
```{r rq1 etable to latex}
# --- RQ1: LaTeX table with fixest::etable ---------------------------------

fixest::etable(list(m_rq1_base, m_rq1_trend, m_rq1_total, m_rq1_noren),
               keep = c("cpi10_id_sd", "lgdppc_c", "ind_c", "ren_c"),
               title      = "RQ1 — Effect of Carbon Pricing on $\\log(\\text{CO}_2)$ Emissions: TWFE Estimates",
               label      = "tab:rq1_twfe",
               notes = "All models include country and year fixed effects. The dependent variable is $\\log(\\text{CO}_2\\text{ per capita})$, except in the 'Total emissions' column where it is $\\log(\\text{CO}_2)$ (population added as a control). CPI is residualized on year FE and standardized within country (idiosyncratic CPI). Standard errors are Driscoll--Kraay (bandwidth $L{=}2$)."
              # , tex= TRUE #uncomment to export as LaTeX
               )

```



## RQ2: Dynamic effects of CPI (lags, Almon)

```{r}
# --- Distributed Lag Model (0–2) ---------------------------------------
m_rq2_dl <- feols(
  log_co2_pc ~ cpi10_id_sd + l(cpi10_id_sd,1) + l(cpi10_id_sd,2) +
               lgdppc_c + ind_c + ren_c | country + year,
  data     = analysis_sample,
  vcov     = "DK",
  panel.id = ~ country + year
)
summary(m_rq2_dl)

# -- - Almon lag (2) -----------------------------------------------
m_rq2_almon2 <- feols(
  log_co2_pc ~ almon(cpi10_id_sd, 2, 2) +
               lgdppc_c + ind_c + ren_c | country + year,
  data     = analysis_sample,
  vcov     = "DK",
  panel.id = ~ country + year
)

summary(m_rq2_almon2)

# --- Total emission

m2_total<- feols(
  log_co2 ~ cpi10_id_sd + l(cpi10_id_sd,1) + l(cpi10_id_sd,2) +
               lgdppc_c + ind_c + ren_c + log_pop | country + year,
  data     = analysis_sample,
  vcov     = "DK",
  panel.id = ~ country + year
)
summary(m2_total)

# --- Wald test for cumulative (δ0 + δ1 + δ2 = 0) -----------------------
wald(m_rq2_dl, "c(cpi10_id_sd, l(cpi10_id_sd,1), l(cpi10_id_sd,2)) = 0")

```

```{r vif}
# 1) Build aligned lag columns
df_vif <- analysis_sample %>%
  arrange(country, year) %>%
  group_by(country) %>%
  mutate(
    cpi_l0 = cpi10_id_sd,
    cpi_l1 = dplyr::lag(cpi10_id_sd, 1),
    cpi_l2 = dplyr::lag(cpi10_id_sd, 2)
  ) %>%
  ungroup() %>%
  # keep only rows where all lags are present
  filter(!is.na(cpi_l0) & !is.na(cpi_l1) & !is.na(cpi_l2))

# 2) Partial-out FE on the *same trimmed sample*
r0 <- fixest::feols(cpi_l0 ~ 1 | country + year, data = df_vif)$residuals
r1 <- fixest::feols(cpi_l1 ~ 1 | country + year, data = df_vif)$residuals
r2 <- fixest::feols(cpi_l2 ~ 1 | country + year, data = df_vif)$residuals

# 3) Now lengths match -> build data.frame and compute VIFs
rx   <- data.frame(r0 = r0, r1 = r1, r2 = r2)
lm_x <- lm(r0 ~ r1 + r2, data = rx)

car::vif(lm_x)            # VIFs for the lag block
kappa(model.matrix(lm_x)) # condition number (large = more collinearity)



```
```{r}
# --- cumulative effect (δ0+δ1+δ2) with CLUSTERED-by-country VCOV ----
library(sandwich)

# names of the three lag coefficients:
lag_names <- c("cpi10_id_sd","l(cpi10_id_sd, 1)","l(cpi10_id_sd, 2)")

# 1) Coef vector and clustered VCOV
b <- coef(m_rq2_dl)[lag_names]
Vc_cl <- sandwich::vcovCL(m_rq2_dl, cluster = model.frame(m_rq2_dl)[["country"]])
V <- Vc_cl[lag_names, lag_names]

# 2) Sum, SE, Wald, p-value
sum_b <- sum(b)
se_sum <- sqrt(sum(V))  # since the gradient for sum is a vector of ones
wald_stat <- (sum_b^2) / (se_sum^2)
p_val <- 1 - pchisq(wald_stat, df = 1)
c(sum = sum_b, se = se_sum, wald = wald_stat, p = p_val)

# 3) 95% CI for the sum, then convert to "% for +10 CPI points"
ci_sum <- sum_b + c(-1,1) * 1.96 * se_sum

sd_id <- sd(analysis_sample$cpi10_id, na.rm = TRUE)  # SD of idiosyncratic CPI in original units
k <- 10                                              # +10 CPI points

pct_point <- 100 * sum_b * (k / sd_id)
pct_ci    <- 100 * ci_sum * (k / sd_id)
c(pct_point = pct_point, ci_lo = pct_ci[1], ci_hi = pct_ci[2])

```
```{r}

# Ensure time is numeric & a simple country trend variable exists
analysis_sample <- analysis_sample %>%
  mutate(year = as.integer(year),
         cyear = year - min(year, na.rm = TRUE)) %>%
  arrange(country, year)

# If you use total emissions in spec (3), ensure these exist:
# analysis_sample$log_co2_total <- log(analysis_sample$co2_total)
# analysis_sample$log_pop       <- log(analysis_sample$population)

# ---- Helper: estimate DL(0–2) with given RHS & DV -------------------------
fit_dl <- function(dv, rhs, data) {
  feols(
    as.formula(
      paste0(dv, " ~ cpi10_id_sd + l(cpi10_id_sd,1) + l(cpi10_id_sd,2) + ",
             rhs, " | country + year")
    ),
    data     = data,
    vcov     = "DK",
    panel.id = ~ country + year
  )
}

# ---- Four specifications (match your RQ1) ---------------------------------
m2_base  <- fit_dl("log_co2_pc",    "lgdppc_c + ind_c + ren_c",               analysis_sample)
m2_trend <- fit_dl("log_co2_pc",    "lgdppc_c + ind_c + ren_c + cyear:country", analysis_sample)
#m2_total <- fit_dl("log_co2_total", "lgdppc_c + ind_c + ren_c + log_pop",     analysis_sample)
m2_noren <- fit_dl("log_co2_pc",    "lgdppc_c + ind_c",                        analysis_sample)


mods <- list(
  "Baseline"          = m2_base,
  "+ Country trends"  = m2_trend,
  "Total emissions"   = m2_total,
  "No renewables"     = m2_noren
)

etable(mods,
       #keep = c("cpi10_id_sd", "l(cpi10_id_sd,1)", "l(cpi10_id_sd,2)"),
       title      = "RQ2 — Dynamic Effects of Carbon Pricing on $\\log(\\text{CO}_2)$ Emissions: TWFE Distributed Lag (0--2) Estimates",
       label      = "tab:rq2_dl_twfe",
       notes = "All models include country and year fixed effects. The dependent variable is $\\log(\\text{CO}_2\\text{ per capita})$, except in the 'Total emissions' column where it is $\\log(\\text{CO}_2)$ (population added as a control). CPI is residualized on year FE and standardized within country (idiosyncratic CPI). Standard errors are Driscoll--Kraay (bandwidth $L{=}2$)."
     #, tex= TRUE #uncomment to export as LaTeX
)
lag_names <- c("cpi10_id_sd", "l(cpi10_id_sd, 1)", "l(cpi10_id_sd, 2)")
sd_id <- sd(analysis_sample$cpi10_id, na.rm = TRUE)  # SD of idiosyncratic CPI in original units
k_pts <- 10

# Wald & percent-effect extractor (clustered-by-country, robust to DK fragility)
summarize_mod <- function(lbl, m) {
  # coefficients for 0,1,2
  b <- coef(m)[lag_names]
  # clustered VCOV
  Vc <- sandwich::vcovCL(m, cluster = model.frame(m)[["country"]])
  V  <- Vc[lag_names, lag_names]

  sum_b  <- sum(b)
  se_sum <- sqrt(sum(V))
  wald   <- (sum_b / se_sum)^2
  pval   <- 1 - pchisq(wald, df = 1)
  ci_sum <- sum_b + c(-1,1) * 1.96 * se_sum

  # % effect for +10 CPI points
  pct_point <- 100 * sum_b  * (k_pts / sd_id)
  pct_ci    <- 100 * ci_sum * (k_pts / sd_id)

  data.frame(
    Specification = lbl,
    delta0   = unname(b[1]),
    delta1   = unname(b[2]),
    delta2   = unname(b[3]),
    SumDL    = unname(sum_b),
    SEsum    = unname(se_sum),
    Wald     = unname(wald),
    p        = unname(pval),
    N        = as.integer(fixest::fitstat(m, "n")),
    Within_R2= as.numeric(fixest::fitstat(m, "wr2")),
    Pct10    = pct_point,
    Pct10_lo = pct_ci[1],
    Pct10_hi = pct_ci[2],
    check.names = FALSE
  )
}

res <- do.call(rbind, Map(summarize_mod, names(mods), mods))

# Nicely formatted table for LaTeX
res_print <- res %>%
  mutate(
    delta0 = sprintf("%.4f", delta0),
    delta1 = sprintf("%.4f", delta1),
    delta2 = sprintf("%.4f", delta2),
    SumDL  = sprintf("%.4f", SumDL),
    SEsum  = sprintf("(%.4f)", SEsum),
    Wald   = sprintf("%.3f", Wald),
    p      = sprintf("%.3f", p),
    Within_R2 = sprintf("%.3f", Within_R2),
    Pct10    = sprintf("%.2f", Pct10),
    Pct10_lo = sprintf("%.2f", Pct10_lo),
    Pct10_hi = sprintf("%.2f", Pct10_hi)
  )

# Save LaTeX to file for your WP
if (!dir.exists("tables")) dir.create("tables", recursive = TRUE)

kbl(
  res_print[, c("Specification","delta0","delta1","delta2","SumDL","SEsum","Wald","p","N","Within_R2","Pct10","Pct10_lo","Pct10_hi")],
  format   = "latex", booktabs = TRUE,
  caption  = "RQ2 — Dynamic TWFE (DL 0–2): lag coefficients, cumulative effect (clustered by country), and percent change for +10 CPI points.",
  label    = "tab:rq2_dl",
  align    = "lrrrrrrrrrrrr"
) %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  add_header_above(c(" " = 1,
                     "Lag coefficients" = 3,
                     "Cumulative effect" = 3,
                     "Fit" = 2,
                     "% effect (+10 CPI)" = 3)) 

# %>%
#   add_footnote(
#     colnames = F, align = "l",
#     values = c("\\multicolumn{13}{l}{\\footnotesize Notes: Models include country and year FE; Driscoll--Kraay SEs in estimation. Joint tests use country-clustered VCOV to avoid DK fragility with short T and multiple lags. Pct effect converts SumDL to % for a +10 CPI-point change using the SD of the idiosyncratic CPI in original units.}")
#   ) %>%
#   cat(file = "tables/rq2_dl.tex")

# Include in TeX:
# \input{tables/rq2_dl.tex}

```


```{r rq2 almon}
library(dplyr)

# Build Almon S0,S1,S2 over K lags using dplyr::lag by country
build_almon_dplyr <- function(df, var, K, id = "country", time = "year"){
  stopifnot(all(c(id, time, var) %in% names(df)))
  df <- df %>% arrange(.data[[id]], .data[[time]])
  # create lag columns
  for(k in 0:K){
    newnm <- paste0(var, "_L", k)
    if(k == 0){
      df[[newnm]] <- df[[var]]
    } else {
      df <- df %>%
        group_by(.data[[id]]) %>%
        mutate(!!newnm := dplyr::lag(.data[[var]], k)) %>%
        ungroup()
    }
  }
  # S0, S1, S2
  lag_names <- paste0(var, "_L", 0:K)
  df$S0_cpi <- rowSums(df[lag_names], na.rm = FALSE)
  df$S1_cpi <- Reduce(`+`, Map(function(k, nm) k   * df[[nm]], 0:K, lag_names))
  df$S2_cpi <- Reduce(`+`, Map(function(k, nm) k^2 * df[[nm]], 0:K, lag_names))
  df
}

K <- 4
dfA <- build_almon_dplyr(analysis_sample, "cpi10_id_sd", K)
#----- Fitting the Almon models with DK SEs -----

mA_base  <- feols(log_co2_pc    ~ S0_cpi + S1_cpi + S2_cpi + lgdppc_c + ind_c + ren_c | country + year,
                  data = dfA, vcov = "DK", panel.id = ~ country + year)
mA_trend <- feols(log_co2_pc    ~ S0_cpi + S1_cpi + S2_cpi + lgdppc_c + ind_c + ren_c + I(year - min(year)):country | country + year,
                  data = dfA, vcov = "DK", panel.id = ~ country + year)
mA_total <- feols(log_co2 ~ S0_cpi + S1_cpi + S2_cpi + lgdppc_c + ind_c + ren_c + log_pop | country + year,
                  data = dfA, vcov = "DK", panel.id = ~ country + year)
mA_noren <- feols(log_co2_pc    ~ S0_cpi + S1_cpi + S2_cpi + lgdppc_c + ind_c | country + year,
                  data = dfA, vcov = "DK", panel.id = ~ country + year)

etable(mA_base, mA_trend, mA_total, mA_noren,
       title      = "RQ2 — Dynamic Effects of Carbon Pricing on $\\log(\\text{CO}_2)$ Emissions: TWFE Almon (2) Estimates",
       label      = "tab:rq2_almon_twfe",
       notes = "All models include country and year fixed effects. The dependent variable is $\\log(\\text{CO}_2\\text{ per capita})$, except in the 'Total emissions' column where it is $\\log(\\text{CO}_2)$ (population added as a control). CPI is residualized on year FE and standardized within country (idiosyncratic CPI). Standard errors are Driscoll--Kraay (bandwidth $L{=}2$)."
     , tex= TRUE #uncomment to export as LaTeX
)
```





```{r}
library(ggplot2)

# --- manually enter your estimated lag coefficients from RQ2 (DL 0–2) ---
dl_coefs <- data.frame(
  k = 0:2,
  delta = c(0.0097, -0.0177, -0.0043)
)

# --- your Almon(2) coefficients (example: replace with actual estimates) ---
a0 <- -0.0021
a1 <- -0.0055
a2 <- 0.0008

# compute implied lag weights for 0–4 lags
lags <- 0:4
almon <- data.frame(
  k = lags,
  delta = a0 + a1*lags + a2*lags^2
)

# combine for plotting
dl_coefs$Model <- "Distributed-lag (DL 0–2)"
almon$Model   <- "Almon polynomial (0–4, degree 2)"
df_plot <- rbind(dl_coefs, almon)

# --- plot ---
ggplot(df_plot, aes(x = k, y = delta, color = Model, shape = Model)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  scale_color_manual(values = c("#0072B2", "#D55E00")) +
  scale_shape_manual(values = c(16, 17)) +
  labs(
    x = "Lag (years)",
    y = expression(paste("Estimated effect ", delta[k])),
    color = NULL, shape = NULL,
    title = "Comparison of Distributed-Lag and Almon(2) Lag Profiles"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

ggsave("figures/rq2_almon_schematic.pdf", width = 7, height = 4)

```

############## strategy

```{r}
# Residualize each regressor on FE
rx <- data.frame(
  cpi   = feols(cpi10_id ~ 1 | country + year, data=analysis_sample)$residuals,
  lgdpp = feols(lgdppc_c ~ 1 | country + year, data=analysis_sample)$residuals,
  ind   = feols(ind_c     ~ 1 | country + year, data=analysis_sample)$residuals,
  ren   = feols(ren_c     ~ 1 | country + year, data=analysis_sample)$residuals
)

# VIFs among residualized X's:
lm_x <- lm(cpi ~ lgdpp + ind + ren, data = rx)
car::vif(lm_x)                 # VIFs
kappa(model.matrix(lm_x))      # condition number

```
```{r}
# Dynamic TWFE with DK SEs (lags 0–2)
m_rq2_alt <- feols(
  log_co2_pc ~ cpi10_id_sd + l(cpi10_id_sd,1) + l(cpi10_id_sd,2) + lgdppc_c + ind_c + ren_c | country + year,
  data = analysis_sample,
  vcov = "DK",
  panel.id = ~ country + year
)
summary(m_rq2_alt)

m_rq2 <- feols(
  log_co2_pc ~ fdl(cpi10_id_sd, 0:2) + lgdppc_c + ind_c + ren_c | country + year,
  data = analysis_sample,
  vcov = "DK",
  panel.id = ~ country + year
)


# Cumulative (short-run) effect δ0+δ1+δ2 and its CI
wald(m_rq2_alt, "c(cpi10_id_sd, fdl(cpi10_id_sd,0:2)l1, fdl(cpi10_id_sd,0:2)l2) = 0")

# (Nice to show) coefficient plot
iplot(m_rq2, x = "cpi10_id_sd")

```



```{r}
# RQ2: Parsimonious lags + cumulative effect
m_rq2_id <- fixest::feols(
  log_co2_pc ~ l(cpi10_id_sd, 0:2) +lgdppc_c + ind_c + ren_c | country + year,
  data= analysis_sample, 
  vcov="DK",
  panel.id = ~ country + year
)
summary(m_rq2_id)
wald(m_rq2_id, "c(cpi10_id_sd, fdl(cpi10_id_sd,0:2)l1, fdl(cpi10_id_sd,0:2)l2)")


```

```{r}
# Degree-2 Almon over 0–4 lags (adjust if you prefer 0–2)
m_rq2_almon <- fixest::feols(
  log_co2_pc ~ palmon(cpi10_id_sd, 0:4, 2) + lgdppc_c + ind_c + ren_c | country + year,
  data     = analysis_sample,
  vcov     = "DK",
  panel.id = ~ country + year
)
summary(m_rq2_almon)

# Optional: recover implied lag weights
plot(multicollinearity = FALSE); iplot(m_rq2_almon, x = "cpi10_id_sd")

```

