---
title: "wp_carbon"
author: "sara"
date: "2025-10-04"
output: html_document
---

# Abstract
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#setwd("~/Library/CloudStorage/OneDrive-umontpellier.fr/Project 2/carbon analysis")
knitr::opts_knit$set(root.dir = "/Users/saramustafazade/Library/CloudStorage/OneDrive-umontpellier.fr/Project 2/carbon analysis")

```

# Data loading and cleaning

```{r}
# Packages
library(skimr)
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(stringr)
# to pull variables
library(eurostat)
library(countrycode)
library(WDI)
# for regression
library(fixest)
library(broom)
library(fixest)
library(modelsummary)
library(glue)
library(plm)
library(sandwich); library(lmtest)
# Load the data

# read the sheet called "final_sheet"
df <- read_excel("carbon_index_2025.xlsx", sheet = "final_sheet")
co2<- read_excel("IEA_EDGAR_CO2_1970_2023.xlsx", sheet = "TOTALS BY COUNTRY")

# check the data
#head(df)
#head(oecd)
```

Filetiring EU countries from our dataset
```{r}

# vector of EU countries
eu_countries <- c(
  "Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus", "Czech Republic",
  "Denmark", "Estonia", "Finland", "France", "Germany", "Greece", "Hungary",
  "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg", "Malta",
  "Netherlands", "Poland", "Portugal", "Romania", "Slovakia", "Slovenia", 
  "Spain", "Sweden"
)
# filter only EU countries
df_eu <- df %>% 
  filter(ds_country %in% eu_countries)  

co2_eu <- co2 %>%
  filter(Name %in% eu_countries)



# check
#head(df_eu)
#head(co2_eu)
```

```{r}
# create a summary table showing the earliest start year for each mechanism in each country
summary_tbl <- df_eu %>% 
  group_by(ds_country, implemented_mechanisms) %>% 
  summarize(start_year = min(start_year, na.rm = TRUE), .groups = "drop" )

#view the summary table
summary_tbl
```

```{r}
# Preparing CO2 emission data for merging
co2_eu_filtered <- co2_eu %>%
  select(IPCC_annex, C_group_IM24_sh, Country_code_A3, Name, Substance,
         Y_2000:Y_2023)

# Reshaping CO2 data from wide to long format

co2_eu_long <- co2_eu %>%
  pivot_longer(cols = starts_with("Y_"),
               names_to = "year",
               values_to = "co2") %>%
  mutate(year = as.integer(sub("Y_", "", year))) %>%
  filter(year >= 2000 & year <= 2023)

# CO2 table: standardize to the same key ---
co2_key <- co2_eu_long %>%
  filter(Substance == "CO2") %>%
  rename(iso3 = Country_code_A3) %>%
  select(iso3, year, co2)   # keep only what's needed

```

```{r df_eu to long}
# ==============================================================================
# ## Step 1: Pivot the Carbon Pricing Index Data to a Long Format
# ==============================================================================
# We select the country identifiers and all columns that start with "carbon_pricing_index_"

cpi_long <- df_eu %>%
  select(ds_country, iso3, starts_with("carbon_pricing_index_")) %>%
  pivot_longer(
    cols = -c(ds_country, iso3),
    names_to = "year",
    values_to = "carbon_pricing_index"
  ) %>%
  mutate(year = str_remove(year, "carbon_pricing_index_"))%>%
  mutate(year = as.integer(year))
# ==============================================================================
# ## Step 3: Join the Two Long Datasets into a Final Panel
# ==============================================================================

# We join by country and year to bring everything together.
# We also convert 'year' to a numeric type for easier analysis.
panel_df <- co2_key %>%
  full_join(cpi_long, by = c( "iso3", "year")) %>%
  mutate(year = as.numeric(year)) %>%
  arrange(iso3, year)

# --- View the result ---
cat("Data has been successfully transformed into a long panel format:\n")
head(panel_df)


```

### Pulling control variables

1) Pull from WDI

```{r}

# Step 1: correct indicator codes
wdi_vars <- c("NY.GDP.MKTP.KD",   # GDP (constant 2015 US$)
              "SP.POP.TOTL",      # Population
              "NV.IND.TOTL.ZS")   # Industry share of GDP (%)

# Step 2: fetch from World Bank
wdi_raw <- WDI(
  country = "all",
  indicator = wdi_vars,
  start = 2000,
  end = 2025
)
# Step 3: Clean and rename (Corrected version)
wdi_clean <- wdi_raw %>%
  rename(
    iso3 = iso3c,  # Use the existing iso3c column as the identifier
    year = year,
    gdp = NY.GDP.MKTP.KD,
    population = SP.POP.TOTL,
    industry_share = NV.IND.TOTL.ZS
  ) %>%
  select(country, iso3, year, gdp, population, industry_share) # Keep the original country name and the new iso3

# Define EU countries in ISO3 codes
eu_iso3 <- c("AUT","BEL","BGR","HRV","CYP","CZE","DNK","EST","FIN","FRA","DEU",
             "GRC","HUN","IRL","ITA","LVA","LTU","LUX","MLT","NLD","POL","PRT",
             "ROU","SVK","SVN","ESP","SWE")

# Keep only EU countries using the clean 'iso3' column
wdi_eu <- wdi_clean %>%
  filter(iso3 %in% eu_iso3)

# Quick check to confirm
head(wdi_eu)


#############

# 2) Merge with your dataset
merged_controls <- panel_df %>%
  left_join(wdi_eu, by = c("iso3", "year"))


# 3) Create transformed controls
merged_controls <- merged_controls %>%
  mutate(
    log_gdp = log(gdp),
    log_pop = log(population),
    gdp_per_capita = gdp / population,
    log_gdp_pc = log(gdp_per_capita)
  )

```
Pulling renewable energy share from Eurostat.
nrg_ind_ren, is one of Eurostat's main energy indicators.
```{r}

# EU-27 (Greece is "EL" in Eurostat)
eu_geo <- c("AT","BE","BG","HR","CY","CZ","DK","EE","FI","FR","DE","EL","HU",
            "IE","IT","LV","LT","LU","MT","NL","PL","PT","RO","SK","SI","ES","SE")

# Pull raw (cache speeds things up), detect the time col (time or TIME_PERIOD)
renewables_raw <- get_eurostat("nrg_ind_ren", cache = TRUE)  # time can be Date or TIME_PERIOD
time_col <- intersect(c("time","TIME_PERIOD"), names(renewables_raw))[1]

# Collapse to one obs per country–year, keep unit == % (PC)
renewables_eu <- renewables_raw %>%
  filter(geo %in% eu_geo, unit == "PC") %>%
  group_by(geo, !!sym(time_col)) %>%
  summarise(renewable_share_pc = mean(values, na.rm = TRUE), .groups = "drop") %>%
  rename(year_raw = !!sym(time_col)) %>%
  mutate(
    year    = as.integer(format(as.Date(year_raw), "%Y")),
    year    = ifelse(is.na(year), as.integer(year_raw), year),
    # Create the 'iso3' column directly with the correct name
    iso3 = countrycode(geo, "eurostat", "iso3c") 
  ) %>%
  # Select only the columns needed, using the 'iso3' name
  select(iso3, year, renewable_share_pc) %>% 
  arrange(iso3, year)

# (The rest of the script for padding and merging remains the same)

# The final merge will now work perfectly
merged_final <- merged_controls %>%
  left_join(renewables_eu, by = c("iso3", "year")) # Merging by 'iso3'
# Safety checks
stopifnot(!anyDuplicated(renewables_eu[c("iso3","year")]))

# Optional: pad to full 2000–2025 panel
renewables_eu <- renewables_eu %>%
  complete(iso3, year = 2000:2025) %>%
  mutate(renewable_share = renewable_share_pc / 100)  # fraction 0–1 (keep both if you like)

# Merge with your main panel (iso3, year)
merged_final <- merged_controls %>%
  left_join(renewables_eu, by = c("iso3","year"))

# Sanity Checks (if necessery uncomment)

# # Post-merge sanity: duplicates? missingness?
# stopifnot(!anyDuplicated(merged_final[c("iso3","year")]))
# 
# # coverage summary
# renewables_eu %>%
#   group_by(iso3) %>%
#   summarise(n_years = n_distinct(year),
#             min_year = min(year), max_year = max(year), .groups = "drop") %>%
#   arrange(iso3)
# 
# nrow(renewables_eu)        # expect ~27 * 20 = 540 (2004–2023)
# length(unique(renewables_eu$iso3))  # expect 27
# length(unique(renewables_eu$year))     # expect 20


```

```{r}

#  Create ONLY the missing outcome variants (keep your existing log_gdp_pc)
merged_final <- merged_final %>%
  mutate(
    # safe logs (NA if non-positive)
    log_co2     = ifelse(co2 > 0, log(co2), NA_real_),
    co2_pc      = co2 / population,
    log_co2_pc  = ifelse(co2_pc > 0, log(co2_pc), NA_real_),
    emis_int    = co2 / gdp,              # emissions intensity
    # optional: fraction 0–1 if you kept renewable_share_pc earlier
    renewable_share = ifelse(!is.na(renewable_share), renewable_share,
                             ifelse(exists("renewable_share_pc"),
                                    renewable_share_pc/100, NA_real_))
  )

# Define your main estimation sample (overlap years; drop NAs in key vars)
analysis_sample <- merged_final %>%
  filter(year >= 2004, year <= 2023) %>%     # renewables start in 2004
  tidyr::drop_na(log_co2, carbon_pricing_index, log_gdp_pc, industry_share, renewable_share)

# Scales & centering (for interpretability & interactions)
analysis_sample <- analysis_sample %>%
  mutate(
    cpi10   = carbon_pricing_index / 10,            # 10-pt change
    lcpi    = log1p(carbon_pricing_index),          # optional robustness
    ren_c   = scale(renewable_share, scale = FALSE),# mean-centered
    lgdppc_c= scale(log_gdp_pc, scale = FALSE),
    ind_c   = scale(industry_share, scale = FALSE),
    t       = year                                   # for trends
  )

```

# Analysis

## RQ1

```{r}
m_rq1_pc <- feols(
log_co2_pc ~ cpi10 + lgdppc_c + ind_c + ren_c | country + year,
data= analysis_sample,
panel.id = ~ country + year,
vcov= ~ country + year # two-way clustered SE
#vcov = vcovDK(time = ~ year, kernel = "Bartlett", bandwidth = 3)
)

# ----- RQ1 robustness: Totals (add log_pop), same within-controls -----
m_rq1_tot <- feols(
log_co2 ~ cpi10 + log_pop + lgdppc_c + ind_c + ren_c | country + year,
data = analysis_sample,
panel.id = ~ country + year,
vcov = ~ country + year,
# vcov = vcovDK(time = ~ year, kernel = "Bartlett", bandwidth = 3)
)

etable(m_rq1_pc, m_rq1_tot, se.below = TRUE,
       headers = c("Per-capita (main)", "Totals (robustness)"))

```
```{r}

models_rq1 <- list(
  "Per-capita (main)"      = m_rq1_pc,
  "Per-capita + Trends"    = m_rq1_pc_trend,
  "Totals (robustness)"    = m_rq1_tot,
  "No Renewables"          = m_pc_noren
)

msummary(
  models_rq1,
  fmt = 3,
  stars = c('*'=.05,'**'=.01,'***'=.001),
  coef_map = c(
    "cpi10"     = "CPI (+10 €/t)",
    "lgdppc_c"  = "log GDP pc (within)",
    "ind_c"     = "Industry share",
    "ren_c"     = "Renewable share",
    "log_pop"   = "log Population"
  ),
  gof_map = c("n","r.squared.within"),
  notes = "Country and year fixed effects in all models. Standard errors clustered two-ways by country and year. Significance: * p < 0.05, ** p < 0.01, *** p < 0.001.",
  output = "html"
)

```
The coefficients on CPI are negative in all models but not statistically significant, implying no immediate reduction in emissions from higher carbon pricing.

Given the scaling (CPI = +10 €/t), even the largest point estimate (−0.007) implies only about 0.7 % lower per-capita emissions per +10 €/t, economically small and statistically uncertain.

Renewable share is consistently large and negative (−1.1 to −1.7, p < 0.01), confirming that clean-energy expansion explains most of the contemporaneous decarbonization.

Adding country-specific trends absorbs some of this variation but does not change the overall conclusion.

```{r app: country trend}

m_rq1_pc_trend <- feols(
  log_co2_pc ~ cpi10 + lgdppc_c + ind_c + ren_c + i(country, year) | country + year,
  data = analysis_sample, vcov = ~ country + year
)
etable(m_rq1_pc, m_rq1_pc_trend, se.below = TRUE,
       headers = c("No trend", "Country-specific trends"))

```
```{r app: renewables}
m_pc_noren <- feols(
  log_co2_pc ~ cpi10 + lgdppc_c + ind_c | country + year,
  data = analysis_sample, vcov = ~ country + year
)
etable(m_rq1_pc, m_pc_noren, headers = c("With renewables","No renewables"))

```
```{r app: high price}
analysis_sample <- analysis_sample %>% mutate(high_price = as.integer(carbon_pricing_index >= 30))
m_rq1_hp <- feols(
  log_co2_pc ~ cpi10*high_price + lgdppc_c + ind_c + ren_c | country + year,
  data = analysis_sample, vcov = ~ country + year
)
# Marginal effect at high prices:
etable(m_rq1_pc, m_rq1_hp, se.below = TRUE,
       headers = c("All prices","With high-price interaction"))

```
## RQ 2

```{r}
# --- RQ2: Dynamics of CPI (immediate vs lagged) ---

# Short DL (0–2): contemporaneous + 1–2 year lags
m_rq2_dl02 <- feols(
  log_co2_pc ~ l(cpi10, 0:2) + lgdppc_c + ind_c + ren_c | country + year,
  data      = analysis_sample,
  panel.id  = ~ country + year,         # <-- REQUIRED for l()
  vcov      = ~ country + year          # or: vcovDK(time = ~year, bandwidth = 3)
)

# Placebo leads to check pre-trends (should be ~0 if no anticipation)
m_rq2_leads_lags <- feols(
  log_co2_pc ~ f(cpi10, 1:2) + l(cpi10, 0:2) + lgdppc_c + ind_c + ren_c | country + year,
  data      = analysis_sample,
  panel.id  = ~ country + year,
  vcov      = ~ country + year
)

# Cumulative (sum of 0–2 lags) Wald test
#wald(m_rq2_dl02, "l(cpi10,0) + l(cpi10,1) + l(cpi10,2) = 0")

# Nicely formatted table
etable(m_rq2_dl02, m_rq2_leads_lags, se.below = TRUE,
       headers = c("DL(0–2)", "Leads(1–2) + DL(0–2)"))

```

```{r}
models_rq2 <- list(
  "Per-capita (main)"      = m_rq2_dl02,
  "Per-capita + Trends"    = m_rq2_leads_lags
)

msummary(
  models_rq2,
  fmt = 3,
  stars = c('*'=.05,'**'=.01,'***'=.001),
  # #coef_map = c(
  #   "cpi10"     = "CPI (+10 €/t)",
  #   "lgdppc_c"  = "log GDP pc (within)",
  #   "ind_c"     = "Industry share",
  #   "ren_c"     = "Renewable share",
  #   "log_pop"   = "log Population"
  # ),
  # gof_map = c("n","r.squared.within"),
  notes = "Dynamics of CPI (immediate vs lagged). Country and year fixed effects in all models. Standard errors clustered two-ways by country and year. Significance: * p < 0.05, ** p < 0.01, *** p < 0.001.",
  output = "html"
)

```

```{r wald rq2}
# Robust cumulative-effect test for lags 0:k (k >= 0)
cum_wald_auto <- function(mod, var = "cpi10", lags = 0:2) {
  nm <- names(coef(mod))
  want <- character()

  for (k in lags) {
    if (k == 0) {
      # contemporaneous is just the bare var name
      cand <- var
    } else {
      # possible printed names: "l(cpi10,1)" or "l(cpi10, 1)"
      cand <- c(sprintf("l(%s,%d)", var, k),
                sprintf("l(%s, %d)", var, k))
    }
    hit <- cand[cand %in% nm]
    if (length(hit) == 0) {
      stop("Could not find coefficient for lag ", k,
           ". Tried: ", paste(cand, collapse = " | "),
           "\nAvailable names include e.g.: ", paste(head(nm, 10), collapse = ", "), " ...")
    }
    want <- c(want, hit[1])
  }

  b <- coef(mod)[want]
  V <- vcov(mod)[want, want, drop = FALSE]
  L <- matrix(1, nrow = 1, ncol = length(want))

  est <- as.numeric(L %*% b)
  se  <- sqrt(as.numeric(L %*% V %*% t(L)))
  z   <- est / se
  p   <- 2 * pnorm(abs(z), lower.tail = FALSE)
  ci  <- est + c(-1, 1) * 1.96 * se
  pct <- 100 * (exp(est) - 1)     # percent interpretation for log DV

  list(est = est, se = se, z = z, p = p, ci = ci, pct = pct, terms = want)
}

# ---- Apply to your RQ2 models ----
#rq2_sum     <- cum_wald_auto(m_rq2_dl02,        var = "cpi10", lags = 0:2)
#rq2_sum_lds <- cum_wald_auto(m_rq2_leads_lags,  var = "cpi10", lags = 0:2)

#rq2_sum
#rq2_sum_lds
# see below for table of these results

```
```{r rq2 table}
# cumulative test (from your helper)
cw_dl   <- cum_wald_auto(m_rq2_dl02, var = "cpi10", lags = 0:2)
cw_lead <- cum_wald_auto(m_rq2_leads_lags, var = "cpi10", lags = 0:2)

# create a small data frame to add as a custom note line
cum_line <- data.frame(
  Variable = "Cumulative CPI (0–2y)",
  `DL(0–2)`           = sprintf("%.3f (%.3f)", cw_dl$est, cw_dl$se),
  `Leads+DL(0–2)`     = sprintf("%.3f (%.3f)", cw_lead$est, cw_lead$se)
)

# show main etable and append manually below
etable(m_rq2_dl02, m_rq2_leads_lags,
       headers = c("DL(0–2)", "Leads(1–2)+DL(0–2)"),
       se.below = TRUE)

print(cum_line, row.names = FALSE)


```


RQ2 (timing). Distributed-lag estimates over t,t−1,t−2 indicate that the contemporaneous CPI effect is close to zero, while the t−1 coefficient is negative and statistically significant (−0.015, p<0.05); the t−2 coefficient is small and imprecise. Placebo leads are insignificant, suggesting no pre-trends. The cumulative 0–2-year effect of a +10 €/t increase is about −1.0% (DL(0–2): −0.0103, SE 0.0205, p=0.615; Leads+DL: −0.0096, SE 0.0197, p=0.628). These patterns imply that most of the policy impact materializes with a one-year delay, consistent with investment and adoption cycles, while the multi-year average over two years is modest and imprecisely estimated.

```{r}
# Coefficient plot focusing on CPI terms
coefplot(
  m_rq2_leads_lags,
  keep = c("f\\(cpi10,1\\)","f\\(cpi10,2\\)","cpi10","l\\(cpi10,1\\)","l\\(cpi10,2\\)"),
  main = "Dynamic effects of CPI (+10 €/t) on log CO2 per capita",
  xlab = "Semi-elasticity",
  dict = c(
    "f(cpi10,1)" = "Lead 1",
    "f(cpi10,2)" = "Lead 2",
    "cpi10"      = "t",
    "l(cpi10,1)" = "t−1",
    "l(cpi10,2)" = "t−2"
  )
)


```
## RQ 3 

```{r}
# --- Construct HighCarbon indicator (pre-2004–2006 baseline) ---
baseline_intensity <- analysis_sample %>%
  filter(year %in% 2004:2006) %>%
  group_by(country) %>%
  summarise(emis_int_base = mean(emis_int, na.rm = TRUE), .groups = "drop") %>%
  mutate(HighCarbon = ifelse(emis_int_base > median(emis_int_base, na.rm = TRUE), 1, 0))

df_rq3 <- analysis_sample %>%
  left_join(baseline_intensity, by = "country")

# --- Heterogeneity regression ---
m_rq3 <- feols(
  log_co2_pc ~ cpi10 * HighCarbon + lgdppc_c + ind_c + ren_c | country + year,
  data = df_rq3, vcov = ~ country + year
)

etable(m_rq3, se.below = TRUE, headers = "H3: High vs. Low Carbon Intensity")

```

```{r}
# Effects and tests
# High group total effect = cpi10 + cpi10:HighCarbon
wald(m_rq3, "cpi10 + cpi10:HighCarbon = 0")   # high group
wald(m_rq3, "cpi10 = 0")                      # low group
wald(m_rq3, "cpi10:HighCarbon = 0")           # difference

```
Test dynamic heterogeneity (this is the one likely to confirm H3)
```{r}
# Dynamic heterogeneity: interact lags with HighCarbon
m_rq3_dyn <- feols(
  log_co2_pc ~ cpi10*HighCarbon + l(cpi10,1)*HighCarbon + l(cpi10,2)*HighCarbon +
               lgdppc_c + ind_c + ren_c | country + year,
  data = df_rq3,
  panel.id = ~ country + year,
  vcov = ~ country + year
)

# Cumulative (0–2y) effects by group:
# low group: sum of cpi10 + l(cpi10,1) + l(cpi10,2)
wald(m_rq3_dyn, "cpi10 + l(cpi10,1) + l(cpi10,2) = 0")

# high group: add the three interactions
wald(m_rq3_dyn, paste(
  "cpi10 + l(cpi10,1) + l(cpi10,2) +",
  "cpi10:HighCarbon + l(cpi10,1):HighCarbon + l(cpi10,2):HighCarbon = 0"
))

# difference in cumulative effects (High − Low)
wald(m_rq3_dyn, paste(
  "cpi10:HighCarbon + l(cpi10,1):HighCarbon + l(cpi10,2):HighCarbon = 0"
))
etable(m_rq3_dyn, se.below = TRUE,
       headers = "H3 Dynamic: High vs. Low Carbon Intensity")
```
Adding a continuous heterogeneity (more power, cleaner)
```{r}
# Standardize baseline intensity (z-score)
df_rq3 <- df_rq3 |>
  mutate(z_intensity = as.numeric(scale(emis_int_base)))

m_rq3_cont <- feols(
  log_co2_pc ~ cpi10*z_intensity + lgdppc_c + ind_c + ren_c | country + year,
  data = df_rq3, vcov = ~ country + year
)

# Slope at −1 SD, mean, +1 SD:
sd1 <- 1
wald(m_rq3_cont, "cpi10 + (-1)*cpi10:z_intensity = 0")  # low baseline
wald(m_rq3_cont, "cpi10 = 0")                           # mean baseline
wald(m_rq3_cont, "cpi10 + ( 1)*cpi10:z_intensity = 0")  # high baseline

```

```{r}
# Find one term among multiple acceptable name variants
.find_one <- function(nm, candidates) {
  hit <- candidates[candidates %in% nm]
  if (length(hit) == 0) return(NA_character_)
  hit[1]
}

# Build name-robust list of terms for dynamic het. (0–2y)
get_terms_dynhet <- function(nm, var = "cpi10", lags = 0:2, int = "HighCarbon") {
  out_low  <- character(0)
  out_high <- character(0)

  for (k in lags) {
    if (k == 0) {
      # base contemporaneous
      base <- .find_one(nm, var)
      # interaction can be in either order, with/without spaces
      int_cands <- c(
        sprintf("%s:%s", var, int),
        sprintf("%s:%s", int, var)
      )
      inter <- .find_one(nm, int_cands)
    } else {
      # lag name variants with/without spaces
      lag_cands <- c(
        sprintf("l(%s,%d)", var, k),
        sprintf("l(%s, %d)", var, k)
      )
      base <- .find_one(nm, lag_cands)

      int_cands <- c(
        sprintf("l(%s,%d):%s", var, k, int),
        sprintf("l(%s, %d):%s", var, k, int),
        sprintf("%s:l(%s,%d)", int, var, k),
        sprintf("%s:l(%s, %d)", int, var, k)
      )
      inter <- .find_one(nm, int_cands)
    }
    out_low  <- c(out_low,  base)
    out_high <- c(out_high, inter)
  }

  list(low = out_low, high = out_high)
}

# Generic linear-combo (sum) with informative errors
lincombo_sum <- function(mod, terms) {
  nm <- names(coef(mod))
  if (anyNA(terms)) {
    miss_k <- which(is.na(terms))
    stop("Some requested terms were not found in the model:\n",
         paste(sprintf("  - component %d is missing", miss_k), collapse = "\n"),
         "\nInspect names(coef(model)) to see available terms.")
  }
  b <- coef(mod)[terms]
  V <- vcov(mod)[terms, terms, drop = FALSE]
  est <- sum(b)
  se  <- sqrt(sum(V))             # because all weights are 1
  z   <- est/se
  p   <- 2*pnorm(abs(z), lower.tail = FALSE)
  ci  <- est + c(-1,1)*1.96*se
  list(est=est, se=se, z=z, p=p, ci=ci, terms=terms)
}


```

```{r}
nm <- names(coef(m_rq3_dyn))
tt <- get_terms_dynhet(nm, var = "cpi10", lags = 0:2, int = "HighCarbon")

# (A) Low-intensity cumulative (0–2y): cpi10 + l1 + l2
res_low  <- lincombo_sum(m_rq3_dyn, tt$low)

# (B) High-intensity cumulative (0–2y): add the three interactions
res_high <- lincombo_sum(m_rq3_dyn, c(tt$low, tt$high))

# (C) Difference in cumulative effects (High − Low): just the three interactions
res_diff <- lincombo_sum(m_rq3_dyn, tt$high)

res_low
res_high
res_diff

```
```{r}
# ---- store your cumulative results ----
rq3_tab <- tribble(
  ~Group,         ~Estimate,    ~`Std. Error`, ~z,        ~`p-value`, ~`95% CI Lower`, ~`95% CI Upper`, ~`% Δ CO2 (cum.)`,
  "Low-intensity", -0.0128,      0.0223,       -0.573,    0.567,      -0.0565,         0.0310,          -1.28,
  "High-intensity",-0.0060,      0.0209,       -0.288,    0.773,      -0.0471,         0.0350,          -0.60,
  "Difference (H–L)", 0.0068,    0.0214,        0.315,    0.753,      -0.0353,         0.0488,           0.68
)

# ---- knit to LaTeX ----
kable(
  rq3_tab, "html", booktabs = TRUE, digits = 4,
  caption = "RQ3: Cumulative 0--2-year effects of Carbon Pricing Index (CPI) by baseline carbon intensity. 
Standard errors clustered two ways (country and year). CPI scaled by +10 €/t."
)


```
Does RQ3 hold with CO2 totals?

```{r}
# Assumes: df_rq3 has HighCarbon and (optionally) z_intensity from before
# If not, rebuild HighCarbon exactly as you did for per-capita.

# --- Dynamic heterogeneity with totals (log CO2) ---
m_rq3_dyn_tot <- feols(
  log_co2 ~ cpi10*HighCarbon + l(cpi10,1)*HighCarbon + l(cpi10,2)*HighCarbon +
            log_pop + lgdppc_c + ind_c + ren_c | country + year,
  data     = df_rq3,
  panel.id = ~ country + year,
  vcov     = ~ country + year
)

# --- Helper to grab correct coefficient names and sum them (as before) ---
.find_one <- function(nm, candidates){ hit <- candidates[candidates %in% nm]; if(!length(hit)) NA_character_ else hit[1] }
get_terms_dynhet <- function(nm, var="cpi10", lags=0:2, int="HighCarbon"){
  low <- high <- character(0)
  for(k in lags){
    if(k==0){
      base <- .find_one(nm, var)
      inter <- .find_one(nm, c(sprintf("%s:%s",var,int), sprintf("%s:%s",int,var)))
    } else {
      lag_cands <- c(sprintf("l(%s,%d)",var,k), sprintf("l(%s, %d)",var,k))
      base <- .find_one(nm, lag_cands)
      int_cands <- c(sprintf("l(%s,%d):%s",var,k,int), sprintf("l(%s, %d):%s",var,k,int),
                     sprintf("%s:l(%s,%d)",int,var,k), sprintf("%s:l(%s, %d)",int,var,k))
      inter <- .find_one(nm, int_cands)
    }
    low  <- c(low,  base)
    high <- c(high, inter)
  }
  list(low=low, high=high)
}
lincombo_sum <- function(mod, terms){
  nm <- names(coef(mod))
  if(anyNA(terms)) stop("Missing terms: ", paste(which(is.na(terms)), collapse=", "),
                        "\nCheck names(coef(model)).")
  b <- coef(mod)[terms]
  V <- vcov(mod)[terms, terms, drop=FALSE]
  est <- sum(b)
  se  <- sqrt(sum(V))     # all weights 1
  z   <- est/se
  p   <- 2*pnorm(abs(z), lower.tail = FALSE)
  ci  <- est + c(-1,1)*1.96*se
  list(est=est, se=se, z=z, p=p, ci=ci, terms=terms)
}

# --- Cumulative (0–2y): Low, High, Difference for totals ---
nm_tot <- names(coef(m_rq3_dyn_tot))
tt_tot <- get_terms_dynhet(nm_tot, var="cpi10", lags=0:2, int="HighCarbon")

res_low_tot  <- lincombo_sum(m_rq3_dyn_tot, tt_tot$low)
res_high_tot <- lincombo_sum(m_rq3_dyn_tot, c(tt_tot$low, tt_tot$high))
res_diff_tot <- lincombo_sum(m_rq3_dyn_tot, tt_tot$high)

res_low_tot
res_high_tot
res_diff_tot

# Optional: put these into an etable GOF block
etable(
  m_rq3_dyn_tot,
  se.below = TRUE,
  headers = "RQ3 (Totals): Dynamic heterogeneity (0–2y)"
  
)
gof = list(
    "Cumulative (Low, 0–2y)"  = sprintf("%.4f (%.4f), p=%.3f", res_low_tot$est,  res_low_tot$se,  res_low_tot$p),
    "Cumulative (High, 0–2y)" = sprintf("%.4f (%.4f), p=%.3f", res_high_tot$est, res_high_tot$se, res_high_tot$p),
    "Diff High−Low (0–2y)"    = sprintf("%.4f (%.4f), p=%.3f", res_diff_tot$est, res_diff_tot$se, res_diff_tot$p)
  )

```
```{r}

# --------------------------
# 0) (Optional) DK vcov (Driscoll–Kraay)
# --------------------------
# Use this if you want DK SEs instead of two-way cluster.
vDK <- function(x) vcovSCC(x, time = ~ year, kernel = "Bartlett", maxlag = 3)

# --------------------------
# 1) Helpers to robustly find names & compute linear combos
# --------------------------
.find_one <- function(nm, candidates){
  hit <- candidates[candidates %in% nm]
  if (!length(hit)) NA_character_ else hit[1]
}

# Collect coefficient names for dynamic heterogeneity: base + lags, and their interactions
get_terms_dynhet <- function(nm, var = "cpi10", lags = 0:2, int = "HighCarbon"){
  low <- high <- character(0)
  for (k in lags) {
    if (k == 0) {
      base   <- .find_one(nm, var)
      inter  <- .find_one(nm, c(sprintf("%s:%s", var, int), sprintf("%s:%s", int, var)))
    } else {
      # handle both l(cpi10,1) and l(cpi10, 1) (with a space)
      lag_cands <- c(sprintf("l(%s,%d)", var, k), sprintf("l(%s, %d)", var, k))
      base      <- .find_one(nm, lag_cands)

      int_cands <- c(sprintf("l(%s,%d):%s", var, k, int),
                     sprintf("l(%s, %d):%s", var, k, int),
                     sprintf("%s:l(%s,%d)", int, var, k),
                     sprintf("%s:l(%s, %d)", int, var, k))
      inter     <- .find_one(nm, int_cands)
    }
    low  <- c(low,  base)
    high <- c(high, inter)
  }
  list(low = low, high = high)
}

# Compute sum(beta_terms) with SE using chosen vcov
lincombo_sum <- function(mod, terms, vcov_arg = NULL){
  nm <- names(coef(mod))
  if (anyNA(terms)) {
    stop("Missing terms at positions: ", paste(which(is.na(terms)), collapse = ", "),
         "\nCheck names(coef(model)) to see how your lags/interactions are labeled.")
  }
  b  <- coef(mod)[terms]
  V  <- if (is.null(vcov_arg)) vcov(mod) else vcov(mod, vcov = vcov_arg)
  V  <- V[terms, terms, drop = FALSE]
  w  <- rep(1, length(terms))
  est <- sum(b)
  se  <- as.numeric(sqrt(t(w) %*% V %*% w))  # = sqrt(sum(V)) since w=1s
  z   <- est / se
  p   <- 2 * pnorm(abs(z), lower.tail = FALSE)
  ci  <- est + c(-1, 1) * 1.96 * se
  list(est = est, se = se, z = z, p = p, ci = ci, terms = terms)
}

# Wrapper to compute Low / High / Diff cumulative (0–2y)
cum_effects_0_2 <- function(mod, var = "cpi10", lags = 0:2, int = "HighCarbon", vcov_arg = NULL){
  nm   <- names(coef(mod))
  tt   <- get_terms_dynhet(nm, var = var, lags = lags, int = int)
  low  <- lincombo_sum(mod, tt$low, vcov_arg = vcov_arg)
  high <- lincombo_sum(mod, c(tt$low, tt$high), vcov_arg = vcov_arg)
  diff <- lincombo_sum(mod, tt$high, vcov_arg = vcov_arg)
  # return a neat data.frame
  tibble::tibble(
    group = c("Low (0–2y)", "High (0–2y)", "High − Low (0–2y)"),
    est   = c(low$est, high$est, diff$est),
    se    = c(low$se,  high$se,  diff$se),
    z     = c(low$z,   high$z,   diff$z),
    p     = c(low$p,   high$p,   diff$p),
    ci_lo = c(low$ci[1], high$ci[1], diff$ci[1]),
    ci_hi = c(low$ci[2], high$ci[2], diff$ci[2])
  )
}

# --------------------------
# 2) Fit BOTH models (PC and TOT) with lags interacted
# --------------------------
# Make sure HighCarbon is 0/1 numeric and data sorted
df_rq3 <- df_rq3 %>%
  mutate(HighCarbon = as.integer(HighCarbon)) %>%
  arrange(country, year)

# Per-capita
m_rq3_dyn_pc <- feols(
  log_co2_pc ~ cpi10*HighCarbon + l(cpi10,1)*HighCarbon + l(cpi10,2)*HighCarbon +
    lgdppc_c + ind_c + ren_c | country + year,
  data     = df_rq3,
  panel.id = ~ country + year,
  vcov     = ~ country + year  # keep two-way cluster here; DK via cum_effects_0_2 if desired
)

# Totals (adds population)
m_rq3_dyn_tot <- feols(
  log_co2 ~ cpi10*HighCarbon + l(cpi10,1)*HighCarbon + l(cpi10,2)*HighCarbon +
    log_pop + lgdppc_c + ind_c + ren_c | country + year,
  data     = df_rq3,
  panel.id = ~ country + year,
  vcov     = ~ country + year
)

# --------------------------
# 3) Cumulative (0–2y) effects using lincombo_sum for BOTH models
# --------------------------
# Use two-way cluster SEs (as estimated) -> vcov_arg = NULL
res_pc_tw  <- cum_effects_0_2(m_rq3_dyn_pc, var = "cpi10", lags = 0:2, int = "HighCarbon", vcov_arg = NULL)
res_tot_tw <- cum_effects_0_2(m_rq3_dyn_tot, var = "cpi10", lags = 0:2, int = "HighCarbon", vcov_arg = NULL)

# If you prefer DK SEs instead, pass vcov_arg = vDK:
# res_pc_dk  <- cum_effects_0_2(m_rq3_dyn_pc,  vcov_arg = vDK)
# res_tot_dk <- cum_effects_0_2(m_rq3_dyn_tot, vcov_arg = vDK)

print(res_pc_tw)
print(res_tot_tw)

# --------------------------
# 4) (Optional) Show models with etable, keep SEs below
# --------------------------
etable(
  m_rq3_dyn_pc, m_rq3_dyn_tot,
  se.below = TRUE,
  headers  = c("RQ3 Dynamic (Per-capita)", "RQ3 Dynamic (Totals + log_pop)")
)

```

```{r}
# Packages
library(modelsummary)
library(dplyr)
library(stringr)
library(tibble)

# -------- Helper: nicely format "est (se)" ----------
fmt_es <- function(est, se, digits = 4) {
  sprintf(paste0("%.", digits, "f (%.", digits, "f)"), est, se)
}

# -------- Build the extra rows for cumulative effects ----------
row_lab_pc  <- "RQ3 Dynamic (Per-capita)"
row_lab_tot <- "RQ3 Dynamic (Totals + log_pop)"

add_rows_tbl <- tibble::tibble(
  term                         = c(" ", "Cumulative (0–2y): Low", "Cumulative (0–2y): High", "Cumulative (0–2y): High − Low"),
  `RQ3 Dynamic (Per-capita)`   = c("", fmt_es(res_pc_tw$est[res_pc_tw$group=="Low (0–2y)"],
                                              res_pc_tw$se [res_pc_tw$group=="Low (0–2y)"]),
                                   fmt_es(res_pc_tw$est[res_pc_tw$group=="High (0–2y)"],
                                              res_pc_tw$se [res_pc_tw$group=="High (0–2y)"]),
                                   fmt_es(res_pc_tw$est[res_pc_tw$group=="High − Low (0–2y)"],
                                              res_pc_tw$se [res_pc_tw$group=="High − Low (0–2y)"])
  ),
  `RQ3 Dynamic (Totals + log_pop)` = c("", fmt_es(res_tot_tw$est[res_tot_tw$group=="Low (0–2y)"],
                                                  res_tot_tw$se [res_tot_tw$group=="Low (0–2y)"]),
                                       fmt_es(res_tot_tw$est[res_tot_tw$group=="High (0–2y)"],
                                                  res_tot_tw$se [res_tot_tw$group=="High (0–2y)"]),
                                       fmt_es(res_tot_tw$est[res_tot_tw$group=="High − Low (0–2y)"],
                                                  res_tot_tw$se [res_tot_tw$group=="High − Low (0–2y)"])
  )
)

# --------- Optional: clean coefficient labels (auto-maps whatever names you have) ----------
nm_pc  <- names(coef(m_rq3_dyn_pc))
nm_tot <- names(coef(m_rq3_dyn_tot))
nm_all <- unique(c(nm_pc, nm_tot))

# Create a mapping that prints nice labels if it recognizes common patterns
pretty_map <- setNames(nm_all, nm_all)
pretty_map[grepl("^cpi10$", nm_all)]                    <- "CPI (t) / +10 €/t"
pretty_map[grepl("^l\\(cpi10,\\s*1\\)$", nm_all)]       <- "CPI (t−1)"
pretty_map[grepl("^l\\(cpi10,\\s*2\\)$", nm_all)]       <- "CPI (t−2)"
pretty_map[grepl("^cpi10:HighCarbon", nm_all)]          <- "× HighCarbon"
pretty_map[grepl("^HighCarbon:cpi10", nm_all)]          <- "× HighCarbon"
pretty_map[grepl("^l\\(cpi10,\\s*1\\):HighCarbon", nm_all)] <- "CPI (t−1) × HighCarbon"
pretty_map[grepl("^HighCarbon:l\\(cpi10,\\s*1\\)", nm_all)] <- "CPI (t−1) × HighCarbon"
pretty_map[grepl("^l\\(cpi10,\\s*2\\):HighCarbon", nm_all)] <- "CPI (t−2) × HighCarbon"
pretty_map[grepl("^HighCarbon:l\\(cpi10,\\s*2\\)", nm_all)] <- "CPI (t−2) × HighCarbon"

pretty_map[grepl("^lgdppc_c$", nm_all)] <- "log GDP pc (centered)"
pretty_map[grepl("^ind_c$", nm_all)]    <- "Industry share (centered)"
pretty_map[grepl("^ren_c$", nm_all)]    <- "Renewables (centered)"
pretty_map[grepl("^log_pop$", nm_all)]  <- "log Population"

# --------- Produce LaTeX table ----------
msummary(
  list(`RQ3 Dynamic (Per-capita)` = m_rq3_dyn_pc,
       `RQ3 Dynamic (Totals + log_pop)` = m_rq3_dyn_tot),
  output   = "latex",
  stars    = c('*' = .05, '**' = .01, '***' = .001),
  coef_map = pretty_map,
  gof_omit = 'IC|Log.Lik|Adj|AIC|BIC|RMSE',      # keep table tight
  add_rows = add_rows_tbl,
  notes = c(
    "+1 in CPI regressor = +10 €/t.",
    "Country & year FE in all models.",
    "SEs: two-way clustered by country & year; cumulative rows show est (SE)."
  ),
  title = "RQ3: Dynamic heterogeneity (0–2 years) across High/Low carbon-intensity countries"
)

```

```{r}
library(kableExtra)

cum_tab <- dplyr::tibble(
  `Cumulative (0–2y)` = c("Low", "High", "High − Low"),
  `Per-capita`        = c(fmt_es(res_pc_tw$est[1],  res_pc_tw$se[1]),
                          fmt_es(res_pc_tw$est[2],  res_pc_tw$se[2]),
                          fmt_es(res_pc_tw$est[3],  res_pc_tw$se[3])),
  `Totals + log_pop`  = c(fmt_es(res_tot_tw$est[1], res_tot_tw$se[1]),
                          fmt_es(res_tot_tw$est[2], res_tot_tw$se[2]),
                          fmt_es(res_tot_tw$est[3], res_tot_tw$se[3]))
)

kbl(cum_tab, format = "html", booktabs = TRUE, align = "lcc",
    caption = "RQ3: Cumulative CPI effects (0–2y) by carbon-intensity group") |>
  kable_classic(full_width = FALSE, html_font = NULL) |>
  add_header_above(c(" " = 1, "Models" = 2)) |>
  footnote(
    general = "+1 in CPI regressor = +10 €/t. Country & year FE. SEs in parentheses.",
    threeparttable = TRUE
  )

```


## RQ 4

```{r}
library(dplyr)
library(fixest)
library(modelsummary)
library(kableExtra)
library(tibble)

# =========================================================
# 0) Construct HighIncome (pre-period 2004–2006)
#    Use an UN-centered GDPpc log variable (e.g., "lgdppc").
# =========================================================
income_var <- "lgdppc_c"   # <-- change if yours is named differently

baseline_income <- analysis_sample %>%
  filter(year %in% 2004:2006) %>%
  group_by(country) %>%
  summarise(lgdppc_base = mean(.data[[income_var]], na.rm = TRUE), .groups = "drop") %>%
  mutate(HighIncome = ifelse(lgdppc_base > median(lgdppc_base, na.rm = TRUE), 1L, 0L))

df_rq4 <- analysis_sample %>%
  left_join(baseline_income, by = "country") %>%
  mutate(HighIncome = as.integer(HighIncome)) %>%
  arrange(country, year)

# =========================================================
# 1) DIRECT EFFECT (no lags), per-capita
# =========================================================
m_rq4_direct_pc <- feols(
  log_co2_pc ~ cpi10 * HighIncome + lgdppc_c + ind_c + ren_c | country + year,
  data = df_rq4,
  vcov = ~ country + year
)

# Optional: direct effect robustness on totals (adds population)
m_rq4_direct_tot <- feols(
  log_co2 ~ cpi10 * HighIncome + log_pop + lgdppc_c + ind_c + ren_c | country + year,
  data = df_rq4,
  vcov = ~ country + year
)

etable(m_rq4_direct_pc, m_rq4_direct_tot, se.below = TRUE,
       headers = c("RQ4 Direct: Per-capita", "RQ4 Direct: Totals + log_pop"))

# =========================================================
# 2) DYNAMIC EFFECT (lags up to 2), per-capita
# =========================================================
m_rq4_dyn_pc <- feols(
  log_co2_pc ~ cpi10*HighIncome + l(cpi10,1)*HighIncome + l(cpi10,2)*HighIncome +
    lgdppc_c + ind_c + ren_c | country + year,
  data     = df_rq4,
  panel.id = ~ country + year,
  vcov     = ~ country + year
)

# 3) DYNAMIC EFFECT robustness on totals
m_rq4_dyn_tot <- feols(
  log_co2 ~ cpi10*HighIncome + l(cpi10,1)*HighIncome + l(cpi10,2)*HighIncome +
    log_pop + lgdppc_c + ind_c + ren_c | country + year,
  data     = df_rq4,
  panel.id = ~ country + year,
  vcov     = ~ country + year
)

etable(m_rq4_dyn_pc, m_rq4_dyn_tot, se.below = TRUE,
       headers = c("RQ4 Dynamic: Per-capita", "RQ4 Dynamic: Totals + log_pop"))

# =========================================================
# 4) CUMULATIVE (0–2y) EFFECTS — helper functions (no wald())
# =========================================================
.find_one <- function(nm, candidates){
  hit <- candidates[candidates %in% nm]
  if (!length(hit)) NA_character_ else hit[1]
}

get_terms_dynhet <- function(nm, var = "cpi10", lags = 0:2, int = "HighIncome"){
  low <- high <- character(0)
  for (k in lags) {
    if (k == 0) {
      base  <- .find_one(nm, var)
      inter <- .find_one(nm, c(sprintf("%s:%s", var, int), sprintf("%s:%s", int, var)))
    } else {
      lag_cands <- c(sprintf("l(%s,%d)", var, k), sprintf("l(%s, %d)", var, k))
      base      <- .find_one(nm, lag_cands)
      int_cands <- c(sprintf("l(%s,%d):%s", var, k, int), sprintf("l(%s, %d):%s", var, k, int),
                     sprintf("%s:l(%s,%d)", int, var, k), sprintf("%s:l(%s, %d)", int, var, k))
      inter     <- .find_one(nm, int_cands)
    }
    low  <- c(low,  base)
    high <- c(high, inter)
  }
  list(low = low, high = high)
}

lincombo_sum <- function(mod, terms, vcov_arg = NULL){
  nm <- names(coef(mod))
  if (anyNA(terms)) stop("Missing terms: ", paste(which(is.na(terms)), collapse = ", "),
                         "\nCheck names(coef(model)).")
  b <- coef(mod)[terms]
  V <- if (is.null(vcov_arg)) vcov(mod) else vcov(mod, vcov = vcov_arg)
  V <- V[terms, terms, drop = FALSE]
  w <- rep(1, length(terms))
  est <- sum(b)
  se  <- as.numeric(sqrt(t(w) %*% V %*% w))
  z   <- est / se
  p   <- 2 * pnorm(abs(z), lower.tail = FALSE)
  ci  <- est + c(-1, 1) * 1.96 * se
  list(est = est, se = se, z = z, p = p, ci = ci, terms = terms)
}

cum_effects_0_2 <- function(mod, var = "cpi10", lags = 0:2, int = "HighIncome", vcov_arg = NULL){
  nm <- names(coef(mod))
  tt <- get_terms_dynhet(nm, var = var, lags = lags, int = int)
  low  <- lincombo_sum(mod, tt$low, vcov_arg = vcov_arg)
  high <- lincombo_sum(mod, c(tt$low, tt$high), vcov_arg = vcov_arg)
  diff <- lincombo_sum(mod, tt$high, vcov_arg = vcov_arg)
  tibble(
    group = c("Low (0–2y)", "High (0–2y)", "High − Low (0–2y)"),
    est   = c(low$est, high$est, diff$est),
    se    = c(low$se,  high$se,  diff$se),
    z     = c(low$z,   high$z,   diff$z),
    p     = c(low$p,   high$p,   diff$p),
    ci_lo = c(low$ci[1], high$ci[1], diff$ci[1]),
    ci_hi = c(low$ci[2], high$ci[2], diff$ci[2])
  )
}

# Compute cumulative effects (two-way clustered SEs from the models)
res_rq4_pc  <- cum_effects_0_2(m_rq4_dyn_pc,  var = "cpi10", int = "HighIncome")
res_rq4_tot <- cum_effects_0_2(m_rq4_dyn_tot, var = "cpi10", int = "HighIncome")

res_rq4_pc
res_rq4_tot

```

```{r app RQ4}
fmt_es <- function(est, se, digits = 4) sprintf(paste0("%.",digits,"f (%.",digits,"f)"), est, se)

cum_tab_rq4 <- tibble(
  `Cumulative (0–2y)` = c("Low", "High", "High − Low"),
  `Per-capita`        = c(fmt_es(res_rq4_pc$est[1],  res_rq4_pc$se[1]),
                          fmt_es(res_rq4_pc$est[2],  res_rq4_pc$se[2]),
                          fmt_es(res_rq4_pc$est[3],  res_rq4_pc$se[3])),
  `Totals + log_pop`  = c(fmt_es(res_rq4_tot$est[1], res_rq4_tot$se[1]),
                          fmt_es(res_rq4_tot$est[2], res_rq4_tot$se[2]),
                          fmt_es(res_rq4_tot$est[3], res_rq4_tot$se[3]))
)

kbl(cum_tab_rq4, format = "html", booktabs = TRUE, align = "lcc",
    caption = "RQ4: Cumulative CPI effects (0–2y) by income group") |>
  kable_classic(full_width = FALSE) |>
  add_header_above(c(" " = 1, "Models" = 2)) |>
  footnote(
    general = "+1 in CPI regressor = +10 €/t. Country & year FE. SEs: two-way clustered by country & year. Values are estimates with SE in parentheses.",
    threeparttable = TRUE
  )

```
Direct effect: per-capita (main) + totals (robustness)

```{r}
library(modelsummary)
library(tibble)
library(stringr)

# Pretty labels for key regressors
pretty_map <- c(
  "cpi10"                              = "CPI (t) / +10 €/t",
  "HighIncome"                         = "HighIncome (baseline absorbed)", # FE absorbs its main effect
  "cpi10:HighIncome"                   = "CPI (t) × HighIncome",
  "lgdppc_c"                           = "log GDP pc (centered)",
  "ind_c"                              = "Industry share (centered)",
  "ren_c"                              = "Renewables (centered)",
  "log_pop"                            = "log Population"
)

# Show clean GOF; hide IC etc.
gof_omit <- "IC|Log.Lik|Adj|AIC|BIC|RMSE"

# Build LaTeX table
dir.create("tables", showWarnings = FALSE)

msummary(
  list(
    "RQ4 Direct: Per-capita (main)"    = m_rq4_direct_pc,
    "RQ4 Direct: Totals + log_pop"     = m_rq4_direct_tot
  ),
  output   = "tables/rq4_direct_main.tex",
  stars    = c('*' = .05, '**' = .01, '***' = .001),
  statistic = "({std.error})",
  coef_map = pretty_map,
  gof_omit = gof_omit,
  notes = c(
    "Country & year fixed effects in all models.",
    "Standard errors two-way clustered by country and year.",
    "+1 in CPI regressor = +10 €/t."
  ),
  title = "RQ4 — Income Heterogeneity: Direct Effect (Per-capita main; Totals robustness)"
)

```

Dynamic effect (lags up to 2): per-capita (main) + totals (robustness)  
```{r}
# Extend pretty labels for lagged terms
pretty_map_dyn <- c(pretty_map,
  "l(cpi10,1)"                         = "CPI (t−1)",
  "l(cpi10,2)"                         = "CPI (t−2)",
  # Interactions (order can vary; map both directions defensively)
  "l(cpi10,1):HighIncome"              = "CPI (t−1) × HighIncome",
  "HighIncome:l(cpi10,1)"              = "CPI (t−1) × HighIncome",
  "l(cpi10,2):HighIncome"              = "CPI (t−2) × HighIncome",
  "HighIncome:l(cpi10,2)"              = "CPI (t−2) × HighIncome"
)

msummary(
  list(
    "RQ4 Dynamic: Per-capita (main)"   = m_rq4_dyn_pc,
    "RQ4 Dynamic: Totals + log_pop"    = m_rq4_dyn_tot
  ),
  output   = "tables/rq4_dynamic_main.tex",
  stars    = c('*' = .05, '**' = .01, '***' = .001),
  statistic = "({std.error})",
  coef_map = pretty_map_dyn,
  gof_omit = gof_omit,
  notes = c(
    "Country & year fixed effects in all models.",
    "Standard errors two-way clustered by country and year.",
    "Dynamic specs include CPI(t), CPI(t−1), CPI(t−2) interacted with HighIncome.",
    "+1 in CPI regressor = +10 €/t."
  ),
  title = "RQ4 — Income Heterogeneity: Dynamic Effects (Per-capita main; Totals robustness)"
)

```

